\section{Alternative hash-based proof of Lemma 4.3}\label{app:restore-by-sampling-hash}

This appendix gives a hash-based realization of the sampling step in
Lemma~4.3.
In the proof from the main text, the sampled family $\{S_{i,j}\}$ is obtained
by explicit independent vertex sampling and stored explicitly.
Here we instead define these sets as threshold sets of pairwise-independent
hash functions, so membership in $S_{i,j}$ can be tested in $O(1)$ time from a
short seed description and the family itself is represented by shared hash
seeds rather than by explicit vertex tables.
This does not change the abstract black-box statement of Lemma~4.3, whose
quantitative bounds remain those proved for the explicit-sampling reduction.
The gain is a sharper accounting of the random family itself: the
explicit-sampling variant contributes an additional
$O\!\left(|G|\log|G|\log\frac{1}{\eps}\right)$ storage term for the sampled
family, whereas the present realization represents that family using only
$O\!\left(\log\frac{1}{\eps}\right)$ shared machine words.
This matters concretely for the distance-1 domination constructions of
Section~6: if the explicit-sampling proof of Lemma~4.3 is replaced there by the
present hash-based one, then the sampled family no longer needs to be stored
explicitly.
Hence the explicit
$O\!\left(|G|\log|G|\log\frac{1}{\eps}\right)$ incidence-table overhead
disappears from the implementation.
This does not change the end-to-end asymptotic class of those constructions:
Section~6 still maintains $O\!\left(\log|G|\log\frac{1}{\eps}\right)$
instances of the relevant weighted structures, and each such instance uses
$O(|G|)$ space there, so the overall space bound remains
$O\!\left(|G|\log|G|\log\frac{1}{\eps}\right)$.

\begin{proof}
Fix an error parameter $\eps\in(0,1)$.
Let $n := |V(G)|$.
We assume that $V(G)=\{1,2,\dots,n\}$, i.e., we identify vertices of $G$ with
integers in $\{1,\dots,n\}$.
In particular, we use the same symbol $v$ both for a vertex and for its
identifier in $\{1,\dots,n\}$.
Set $L := \lfloor \log_2 n\rfloor + 1$ and
\[
\lambda
:= \left\lceil 10 \cdot
\ln\!\left(\tfrac{1}{\eps}\right)\right\rceil.
\]
The ceiling ensures that $\lambda\ge 1$ even when $\eps$ is close to~$1$.
We use the slightly larger constant $10$ here, rather than the constant $8$
from the explicit-sampling proof, only to streamline the estimates below.
The guarantee holds against an oblivious adversary, meaning that the entire
update sequence is fixed independently of the random bits chosen at
initialization.
In particular, for any fixed time step the resulting graph state~$G$ is
independent of the random choices made below.

\smallskip
\noindent\textbf{Pairwise-independent hashing.}
Let $p$ be a prime such that $8n < p < 16n$
(such a prime exists by Bertrand's postulate),
and fix any such prime $p$ at initialization.
For each $j\in\{1,\dots,\lambda\}$ choose $a_j,b_j\in\{0,1,\dots,p-1\}$
independently and uniformly at random and define
$h_j(v) := \bigl(a_j\cdot v+b_j\bigr)\bmod p$.
For every vertex $v$ and $x\in\{0,\dots,p-1\}$, for each choice of $a_j$ there
is exactly one choice of $b_j$ with $h_j(v)=x$, so $h_j(v)$ is uniform.
For any two distinct vertices $u,v$ and any $(x,y)\in\{0,\dots,p-1\}^2$, the
system of congruences $a_j u+b_j\equiv x\pmod p$ and
$a_j v+b_j\equiv y\pmod p$ has a unique solution
$a_j\equiv (x-y)\cdot (u-v)^{-1}\pmod p$ and $b_j\equiv x-a_j u\pmod p$.
Thus $(h_j(u),h_j(v))$ is uniform over $\{0,\dots,p-1\}^2$, and hence each $h_j$
is pairwise-independent.
We use the standard word-RAM model with word size $\Omega(\log n)$; since
$p<16n$, each number in $\{0,\dots,p-1\}$ fits in $O(1)$ words and arithmetic
modulo $p$ takes $O(1)$ time.
The prime $p$ together with all seeds $(a_j,b_j)$ occupies $O(\lambda)$
machine words.

\smallskip
\noindent\textbf{Random sampling family and the maintained sums.}
For every $i\in\{0,1,\dots,L\}$ and $j\in\{1,2,\dots,\lambda\}$ we define
the threshold
\[
\tau_i := \left\lfloor \frac{p}{2^{i+1}} \right\rfloor
\]
and the set
\[
S_{i,j} := \bigl\{v\in V(G) : h_j(v) < \tau_i\bigr\}.
\]
Since $L=\lfloor\log_2 n\rfloor+1$, we have $n<2^{L}\le 2n$,
and consequently $2^{L+1}\le 4n$.
Therefore, for every $i\in\{0,\dots,L\}$ we have
$2^{i+1}\le 2^{L+1}\le 4n$.
Since $p>8n$, this implies $p/2^{i+1} > 8n/(4n)=2$ and hence
$\tau_i=\lfloor p/2^{i+1}\rfloor \ge 2$.
Define $p_i := \tau_i/p$.
Because $h_j(v)$ is uniform over $\{0,\dots,p-1\}$,
for every vertex~$v$ we have
\[
\Pr(v\in S_{i,j})
= \frac{\tau_i}{p}
= p_i,
\]
and for every pair of distinct vertices $u,v\in V(G)$
(by the pairwise-independence property of~$h_j$),
\begin{equation}\label{eq:pairwise}
\Pr(u\in S_{i,j} \;\wedge\; v\in S_{i,j})
= p_i^{\,2}.
\end{equation}
The random choices are performed once at initialization and remain fixed
afterwards.
Crucially, membership $v\in S_{i,j}$ can be tested in $O(1)$ time by
evaluating $h_j(v)$ from the stored seed; the sets $S_{i,j}$ are
\emph{not} stored explicitly.

For every $(i,j)$ define two weighting functions
$w^{(0)}_{i,j},w^{(1)}_{i,j}\colon V(G)\to\Z$ by
\[
w^{(0)}_{i,j}(v) :=
\begin{cases}
1 & \text{if } v\in S_{i,j},\\
0 & \text{otherwise,}
\end{cases}
\qquad
w^{(1)}_{i,j}(v) :=
\begin{cases}
v & \text{if } v\in S_{i,j},\\
0 & \text{otherwise.}
\end{cases}
\]
Once the hash seeds are fixed, these are ordinary fixed weighting functions
$w\colon V(G)\to\Z$, so they are legitimate inputs for the correctness
argument below.
Moreover, each value $w^{(\ell)}_{i,j}(v)$ can be evaluated in $O(1)$ time from
the seed of~$h_j$ and the index~$i$, and the description of the weighting
function itself uses only $O(1)$ machine words.
Each maintained instance therefore needs only its own constant-size local
descriptor of $(i,j,\ell)$ in addition to the shared hash seeds.
This compact representation is specific to the present implementation and is
\emph{not} part of the abstract hypothesis of Lemma~4.3.
Accordingly, the proof below uses only that the maintained instances correspond
to fixed weighting functions.
Any quantitative gain from the hash-based representation requires a concrete
implementation of $\mathsf{WeightedSum}_{f,w}(G)$ that can work with such
implicitly described weights without materializing the full table $w$.
This is exactly the setting of the Section~6 distance-1 domination
instantiations, whose bounds are stated besides the read-only representation of
$w$.

We maintain, for every $(i,j)$, two instances of the promised structure:
$\mathsf{WeightedSum}_{f,\,w^{(0)}_{i,j}}(G)$ and
$\mathsf{WeightedSum}_{f,\,w^{(1)}_{i,j}}(G)$.
In \cref{it:general-f} we additionally maintain one instance of
$\mathsf{Verify}_{f}(G)$.
Every update to~$G$ is relayed to all maintained instances.

\smallskip
\noindent\textbf{Query algorithm.}
Given the current graph~$G$, for all pairs
$(i,j)\in\{0,\dots,L\}\times\{1,\dots,\lambda\}$ in lexicographic order, let
\[
A_{i,j} := \sum_{v\in V(G)} f(G,v)\,w^{(0)}_{i,j}(v),
\qquad
B_{i,j} := \sum_{v\in V(G)} f(G,v)\,w^{(1)}_{i,j}(v),
\]
obtained by querying the respective maintained instances.
If $A_{i,j}=0$, continue to the next pair.
\begin{itemize}
\item In \cref{it:binary-f}, if $A_{i,j}=1$, return the vertex
$B_{i,j}$ (we verify below that then $B_{i,j}\in\{1,\dots,n\}$); otherwise,
continue.
\item In \cref{it:general-f}, since $A_{i,j}>0$, compute $q := B_{i,j}/A_{i,j}$.
If $q$ is not an integer or $q\notin\{1,\dots,n\}$, continue.
Otherwise, query $\mathsf{Verify}_f(G)$ on the vertex $q$;
if $f(G,q)>0$, return $q$, and otherwise continue.
\end{itemize}
If no pair leads to returning a vertex, output the distinguished answer
\textsc{none}.

\smallskip
\noindent\textbf{No false positives.}
In \cref{it:binary-f}, suppose the algorithm returns $B_{i,j}$ for
some $(i,j)$.
By definition of $w^{(0)}_{i,j}$ and the assumption
$f(G,v)\in\{0,1\}$, we have
\[
A_{i,j}=\sum_{v\in S_{i,j}} f(G,v)=1,
\]
hence there exists exactly one vertex $x\in S_{i,j}$ with $f(G,x)=1$, and
for all $v\in S_{i,j}\setminus\{x\}$ we have $f(G,v)=0$.
Therefore
\[
B_{i,j}=\sum_{v\in S_{i,j}} f(G,v)\,v=x,
\]
so the returned vertex is $B_{i,j}=x$ and satisfies
$f(G,x)=1>0$.

In \cref{it:general-f}, the algorithm returns a vertex $q$ only after
$\mathsf{Verify}_f(G)$ has confirmed that $f(G,q)>0$; hence every
returned vertex satisfies $f(G,v)>0$.
In particular, even if $q$ happens to be an integer in $\{1,\dots,n\}$ but
corresponds to a vertex $y$ with $f(G,y)=0$, the verification step prevents
returning~$y$.

If the algorithm outputs \textsc{none}, then it makes no positive claim.
Thus, in both cases the data structure has no false positives.

\smallskip
\noindent\textbf{Probability of a false negative.}
Fix a state of the graph~$G$ and let
\[
X := \{v\in V(G)\mid f(G,v)>0\}
\qquad\text{and}\qquad
m := |X|.
\]
Formally, we consider a fixed time step (that is, a graph state obtained after
a fixed number of updates); since the update sequence is fixed independently of
the random bits of the construction, the resulting graph~$G$ (and hence $X$)
is determined independently of the hash functions $h_1,\dots,h_\lambda$.
Equivalently, we may condition on this fixed graph state~$G$ without changing
the distribution of the random hash functions.
If $m=0$, then $f(G,v)=0$ for all $v\in V(G)$ and consequently $A_{i,j}=0$ for
every $(i,j)$, so the query algorithm correctly returns \textsc{none}.
Hence assume $m\ge 1$.

Set $i^\star := \lfloor \log_2 m\rfloor + 1$.
Since $m\le n$ and $L = \lfloor\log_2 n\rfloor + 1$, we have
$i^\star \le L$, so $i^\star\in\{0,1,\dots,L\}$.
Note that for $m=1$ we have $i^\star=1$.
Moreover, $2^{i^\star-1}\le m < 2^{i^\star}$.

Define $\mu := m\cdot p_{i^\star} = m\cdot \tau_{i^\star}/p$.
Since $\tau_{i^\star}=\lfloor p/2^{i^\star+1}\rfloor \le p/2^{i^\star+1}$,
we have
\[
\mu \le m\cdot 2^{-i^\star-1}
< 2^{i^\star}\cdot 2^{-i^\star-1}
= 1/2.
\]
Moreover, $\tau_{i^\star} > p/2^{i^\star+1}-1$, so
\[
\mu > m\cdot 2^{-i^\star-1} - \frac{m}{p}
\ge 2^{i^\star-1}\cdot 2^{-i^\star-1} - \frac{m}{p}
= \frac14 - \frac{m}{p}.
\]
$p>8n\ge 8m$, so $m/p < 1/8$, and therefore
$\mu\in(1/8,\,1/2)$.

Fix any $j\in\{1,\dots,\lambda\}$.
For each $v\in X$ let $I_v := \mathbf{1}[v\in S_{i^\star,j}]$ be the
indicator random variable and set $Y := \sum_{v\in X} I_v
= |S_{i^\star,j}\cap X|$.
We have $\Pr(I_v = 1) = p_{i^\star}$ for each $v\in X$, and by the
pairwise-independence property~\eqref{eq:pairwise}, for any two distinct
vertices $u,v\in X$:
\[
\Pr(I_u = 1 \;\wedge\; I_v = 1) = p_{i^\star}^{\,2}.
\]

We lower-bound $\Pr(Y=1)$.
For each $v\in X$, the event $\{I_v=1\}$ is the disjoint union of the events
\[
E_v := \Bigl\{I_v = 1 \;\wedge\; \bigwedge_{u\in X\setminus\{v\}} I_u=0\Bigr\}
\qquad\text{and}\qquad
F_v := \bigl\{I_v=1\;\wedge\;\exists\,u\in X\setminus\{v\}:\;I_u=1\bigr\},
\]
hence $\Pr(E_v)=\Pr(I_v=1)-\Pr(F_v)$.
Observe that the event $\{Y=1\}$ is exactly the disjoint union
$\biguplus_{v\in X} E_v$, and hence $\Pr(Y=1)=\sum_{v\in X}\Pr(E_v)$.
By the union bound we have
\[
\Pr(F_v) \le \sum_{u\in X\setminus\{v\}} \Pr(I_v=1\;\wedge\;I_u=1).
\]
Therefore:
\begin{align*}
\Pr(Y=1)
&= \sum_{v\in X} \Pr\!\Bigl(I_v=1 \;\wedge\;
\textstyle\bigwedge_{u\in X\setminus\{v\}} I_u=0\Bigr)\\
&= \sum_{v\in X}\Bigl[\Pr(I_v=1)
- \Pr\!\bigl(I_v=1\;\wedge\;\exists\,u\in X\setminus\{v\}:\;I_u=1\bigr)
\Bigr]\\
&\ge \sum_{v\in X}\Bigl[p_{i^\star}
- \sum_{u\in X\setminus\{v\}} \Pr(I_v=1 \;\wedge\; I_u=1)\Bigr]\\
&= m\,p_{i^\star} - m(m-1)\,p_{i^\star}^{\,2}
\;=\; \mu - \tfrac{m-1}{m}\,\mu^2
\;\ge\; \mu - \mu^2
\;=\; \mu(1-\mu).
\end{align*}
The function $g(\mu):=\mu(1-\mu)$ satisfies $g'(\mu)=1-2\mu > 0$ for
$\mu<1/2$, hence $g$ is increasing on $[0,\,1/2)$.
Therefore $g(\mu) > g(1/8) = 7/64 > 1/10$, and so
\[
\Pr\bigl(|S_{i^\star,j}\cap X|=1\bigr)=\Pr(Y=1)\ge \tfrac{1}{10}.
\]

The hash functions $h_1,\dots,h_\lambda$ are sampled independently.
For each fixed $j$, the set $S_{i^\star,j}$ (and hence the event
$\{|S_{i^\star,j}\cap X|\neq 1\}$) is determined solely by~$h_j$.
Therefore, the events $\{|S_{i^\star,j}\cap X|\neq 1\}$ are independent across
distinct values of~$j$.
Consequently
\[
\Pr\Bigl(\forall\, j\in\{1,\dots,\lambda\}\colon\;
|S_{i^\star,j}\cap X|\neq 1\Bigr)
\le \left(1-\tfrac{1}{10}\right)^\lambda
\le \exp\!\left(-\tfrac{\lambda}{10}\right)
\le \exp\!\left(-\ln\!\left(\tfrac{1}{\eps}\right)\right)
= \eps,
\]
where we used $(1-x)\le e^{-x}$ for $x\ge 0$ and the definition
$\lambda\ge 10\ln(1/\eps)$.

We now argue that if for some~$j$ we have $|S_{i^\star,j}\cap X|=1$, then
the query algorithm returns a vertex from~$X$ no later than when it processes
the pair $(i^\star,j)$.
Let~$x$ be the unique element of $S_{i^\star,j}\cap X$.
By definition of $X$, we have $f(G,x)>0$.
\begin{itemize}
\item In \cref{it:binary-f}, $f(G,x)=1$ and $f(G,v)=0$ for all
$v\in S_{i^\star,j}\setminus\{x\}$
(since every such $v$ satisfies $v\notin X$, hence $f(G,v)=0$ by
definition of~$X$), so
$A_{i^\star,j}=\sum_{v\in S_{i^\star,j}}f(G,v)=f(G,x)=1$ and
$B_{i^\star,j}=\sum_{v\in S_{i^\star,j}}f(G,v)\,v=x$.
The query algorithm returns~$x$ when it encounters this pair.
\item In \cref{it:general-f}, since $x$ is the only element of
$S_{i^\star,j}$ with $f(G,\cdot)>0$ (every other $v\in S_{i^\star,j}$
satisfies $v\notin X$ and hence $f(G,v)=0$), we have
$A_{i^\star,j}=f(G,x)>0$ and $B_{i^\star,j}=f(G,x)\cdot x$.
Therefore $q = B_{i^\star,j}/A_{i^\star,j}=x\in\{1,\dots,n\}$,
which is an integer.
The algorithm then queries $\mathsf{Verify}_f(G)$ on~$x=q$,
which confirms $f(G,x)>0$, and returns~$x$.
\end{itemize}
Moreover, if the algorithm returns a vertex for any \emph{earlier} pair
$(i',j')$ --- not necessarily with $i'=i^\star$ --- the returned vertex still
satisfies $f(G,v)>0$ by the no-false-positives property, so this is also
not a false negative.
Hence a false negative occurs only when the algorithm exhausts all pairs
without returning, which requires
$|S_{i^\star,j}\cap X|\neq 1$ for every $j\in\{1,\dots,\lambda\}$, an event
of probability at most~$\eps$.
Thus, for every fixed query asked at any fixed time step, under an oblivious
adversary, the probability that the data structure outputs \textsc{none}
despite $X\neq\emptyset$ is at most $\eps$.
This is the false-negative guarantee relevant for Lemma~4.3.

\smallskip
\noindent\textbf{Complexity.}
We maintain $2(L+1)\lambda$ instances of~$\mathsf{WeightedSum}_{f,w}(G)$.
In \cref{it:general-f} we additionally maintain one instance of
$\mathsf{Verify}_f(G)$.
Since $L=\lfloor\log_2 n\rfloor+1$, we have $(L+1)\lambda =
O\!\left(\log|G|\cdot \log\frac{1}{\eps}\right)$ in the asymptotic style of the
lemma statement.

\emph{The random family itself.}
The hash-based representation stores the random family in
$O\!\left(\log\frac{1}{\eps}\right)$ shared machine words.
Thus the family $\{S_{i,j}\}$ itself occupies only
$O\!\left(\log\frac{1}{\eps}\right)$ space and supports membership tests in
$O(1)$ time, without storing any set $S_{i,j}$ explicitly.
Each maintained instance carries only a constant-size local descriptor of the
corresponding weighting function $w^{(\ell)}_{i,j}$.
This sharper accounting concerns the random family in the present
implementation; it does not alter the conservative abstract statement of
Lemma~4.3.
Its implementation-level effect is visible precisely in instantiations, such as
those of Section~6, where the sampled incidence tables disappear from the
accounting even though the dominant asymptotic term still comes from the many
maintained weighted structures.

\emph{Initialization.}
After fixing the shared hash description, the reduction no longer performs
explicit vertex-by-vertex sampling for every pair $(i,j)$ and does not
materialize the sets $S_{i,j}$.
Initializing all $2(L+1)\lambda$ maintained instances of
$\mathsf{WeightedSum}_{f,w}(G)$ costs
$O\!\bigl((L+1)\lambda\cdot I_R\bigr)$, and in
\cref{it:general-f} initializing the maintained instance of
$\mathsf{Verify}_f(G)$ costs $O(I_V)$.
Hence the total initialization bound is
\[
O\!\left(I_R\log|G|\log\tfrac{1}{\eps} + I_V\right).
\]
Thus the random family contributes only
$O\!\left(\log\tfrac{1}{\eps}\right)$ shared initialization data, rather than
the explicit construction of all sampled sets.
Any further end-to-end initialization gain depends on an implementation of
$\mathsf{WeightedSum}_{f,w}(G)$ that accepts the compact descriptions of
$w^{(0)}_{i,j}$ and $w^{(1)}_{i,j}$ directly.

\emph{Amortized update time.}
Every update to~$G$ is relayed to all $2(L+1)\lambda$ instances of
$\mathsf{WeightedSum}$ and to the single instance of $\mathsf{Verify}$
(in \cref{it:general-f}).
Thus, every update to~$G$ triggers $2(L+1)\lambda$ update operations in
$\mathsf{WeightedSum}$ instances and one update operation in
$\mathsf{Verify}$ (in \cref{it:general-f}).
Summing the amortized bounds over all maintained instances, the total
amortized update time is therefore
\[
O\!\bigl(2T_R\cdot (L+1)\lambda + T_V\bigr)
\subseteq
O\!\bigl((T_R+T_V)\cdot (L+1)\lambda\bigr)
=
O\!\left((T_R+T_V)\cdot \log|G|\cdot
\log\tfrac{1}{\eps}\right).
\]

\emph{Amortized query time.}
One outer query inspects at most $(L+1)\lambda$ pairs.
For each processed pair it performs two queries to maintained instances of
$\mathsf{WeightedSum}_{f,w}(G)$ and, in case \cref{it:general-f}, at most one
query to $\mathsf{Verify}_f(G)$.
Therefore the total black-box query cost of one outer query is
\[
O\!\bigl((Q_R+Q_V)\cdot (L+1)\lambda\bigr).
\]
In \cref{it:general-f}, for each processed pair with $A_{i,j}>0$ the reduction
additionally computes $q=B_{i,j}/A_{i,j}$ and tests integrality and membership
in $\{1,\dots,n\}$.
Under the standard word-RAM convention, these arithmetic operations contribute
only $O(1)$ time per processed pair.
If one wants to charge exact arithmetic on the returned values separately, this
is an additional model-dependent overhead not captured by the black-box bounds
$Q_R,Q_V$.
Hence the amortized query time is
\[
O\!\bigl((Q_R+Q_V)\cdot (L+1)\lambda\bigr)
=
O\!\left((Q_R+Q_V)\cdot \log|G|\cdot
\log\tfrac{1}{\eps}\right).
\]

\emph{Space.}
The $2(L+1)\lambda$ instances of $\mathsf{WeightedSum}$ use
$O\!\bigl(S_R\cdot (L+1)\lambda\bigr)$ space.
The single instance of $\mathsf{Verify}$ uses $O(S_V)$ space.
Thus the maintained data structures contribute
\[
O\!\left(S_R\log|G|\log\tfrac{1}{\eps} + S_V\right)
\]
space in the asymptotic style of the lemma statement.
In addition, the hash-based representation of the random family uses only
$O\!\left(\log\tfrac{1}{\eps}\right)$ shared machine words, plus
constant-size local descriptors for the maintained instances.
Thus the explicit-sampling overhead
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$ for storing the sampled family
is replaced, at the level of the random family itself, by
$O\!\left(\log\tfrac{1}{\eps}\right)$ shared hash data and no explicit
per-vertex tables.
Hence, if one replaces the explicit-sampling proof of Lemma~4.3 by the present
hash-based implementation in the Section~6 distance-1 domination
constructions, then the sampled family is no longer stored explicitly there.
This removes a redundant
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$ block of sampled-incidence
data.
However, the end-to-end asymptotic space bound of the Section~6 constructions
is still dominated by the $2(L+1)\lambda$ maintained
$\mathsf{WeightedSum}$ structures (and the single $\mathsf{Verify}$ structure
in \cref{it:general-f}).
Substituting the Section~6 bounds $S_R=O(|G|)$ and $S_V=O(|G|)$ into the
displayed space formula above therefore still gives
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$ overall space.

In the case \cref{it:binary-f} we adopt the convention
$I_V=T_V=Q_V=S_V=0$, exactly as in the lemma statement.
\end{proof}

As a consequence specific to the distance-1 domination instantiations of
Section~6, replacing the explicit-sampling proof of Lemma~4.3 by the present
hash-based proof removes the explicit storage of the sampled family. In the
explicit-sampling realization, storing the sampled incidence tables contributes
an additional
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$ term. Here, by contrast, the
family is represented by only
$O\!\left(\log\tfrac{1}{\eps}\right)$ shared hash data, so the explicit
per-vertex storage term disappears. The gain is therefore the elimination of a
heavy auxiliary
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$ data structure, rather than a
change in the dominant asymptotic term. Section~6 still maintains
$O\!\left(\log|G|\log\tfrac{1}{\eps}\right)$ instances of
$\mathsf{WeightedSum}$, each using $O(|G|)$ space, and the single maintained
$\mathsf{Verify}$ instance is absorbed in the same bound. Consequently, after
this proof replacement, the Section~6 end-to-end space bound remains
$O\!\left(|G|\log|G|\log\tfrac{1}{\eps}\right)$.
