\section{Counting and finding small patterns in sparse graphs}\label{sec:dvorak}



%Here we recall and extend relevant stuff from Dvo\v{r}\'ak and T\r{u}ma~\cite{DvorakT13}.

%\wninline{Zdaje się, że zapomniałem pospecyfikować w paru miejscach, że złożoności czasowe są amortyzowane}

In this section we first recall the main results of Dvo\v{r}\'ak and T\r{u}ma~\cite{DvorakT13} about dynamic data structures for counting homomorphisms and (induced) subgraphs in classes of bounded expansion, and then we generalize them in various ways for the sake of using them in the next section.
The main result Dvo\v{r}\'ak and T\r{u}ma can be expressed as follows (this is a slightly more general formulation of \cref{thm:overview-DT}).%\michal{Do Dvorak and Tuma have colors? It would be good to cite a specific theorem.} Here, by a \emph{$k$-colored graph} we mean a graph whose edges are colored with the set of colors $[k]$.

\begin{theorem}[\cite{DvorakT13}] \label{thm:dvorak-tuma}
	Let $H$ be a fixed graph, $\G$ be a graph class of bounded expansion, $k\in \N$, and $\F$ be either $\Hom$, $\Sub$, or $\ISub$. Suppose $G$ is a dynamic $k$-colored graph on $n$ vertices that belongs to $\G$ at all times. Then, there is a~data structure $\UnMaps_{\F, H}[G]$ that is able to determine the quantity $|\F(H, G)|$ after every update. The amortized update time is $\Oh_{\G, H, k}(\log^h n)$, where $h = {|H| \choose 2} - 1$, the initialization time is $\Oh_{\G,H,k}(n)$, and the space complexity is $\Oh_{\G, H, k}(n)$. %If $\G$ is a nowhere-dense class, then each update takes $\Oh_{\G, H, k}(|G|^\delta)$ time and its space complexity is $\Oh(|G|^{1 + \delta})$.
\end{theorem}

\subsection{Finding examples of mappings}

%However, they concluded that even though they are able to count appearances of $H$ in $G$ as an induced subgraph (that is, determine $|\ISub(H, G)|$), it is not easy to restore any example of such mapping from their data structure as it involves inclusion-exclusion principle. Despite not written explicitly, it can be expected that the authors would have similar troubles restoring examples of subgraph isomorphisms (as the reduction from counting subgraph isomorphisms to counting homomorphisms requires subtractions as well), though they had likely known how to restore examples of homomorphisms, as that part of the algorithm does not require any subtractions and the restoration of an example can be done through tracing back transitions with nonzero contributions in the designed dynamic programming. 

While the data structure of \cref{thm:dvorak-tuma} is able to \emph{count} the appearances of $H$ in $G$ as an induced subgraph (that is, determine $|\ISub(H, G)|$), it is not easy to restore any example of such a mapping from the data structure, because the counting is performed using the Inclusion--Exclusion Principle. In fact, Dvo\v{r}\'ak and T\r{u}ma explicitly ask the question whether examples of induced subgraph isomorphisms can be also reported efficiently~\cite[Section~5]{DvorakT13}. While the same issue also appears when counting subgraphs, restoring examples of homomorphisms can be done by carefully tracing transitions with nonzero contributions in the designed dynamic programming, as this part of the argumentation in \cite{DvorakT13} does not involve subtractions or counting using the Inclusion-Exclusion Principle.


%However, they concluded that even though they are able to count appearances of $H$ in $G$ as an induced subgraph (that is, determine $|\ISub(H, G)|$), it is not easy to restore any example of such mapping from their data structure as it involves inclusion-exclusion principle. Despite not written explicitly, it can be expected that the authors would have similar troubles restoring examples of subgraph isomorphisms (as the reduction from counting subgraph isomorphisms to counting homomorphisms requires subtractions as well), though they had likely known how to restore examples of homomorphisms, as that part of the algorithm does not require any subtractions and the restoration of an example can be done through tracing transitions with nonzero contributions in the designed dynamic programming. 

In this section, we resolve all these open questions and prove that example mappings can be efficiently reported both for homomorphisms and for (induced) subgraphs, at the cost of making the data structure randomized.
Formally, we prove the following extension of \cref{thm:dvorak-tuma}, already announced in \cref{sec:overview}.

%\wninline{Chyba dobrze by było dodać $\eps$ do parametrów struktury danych, i.e. $\MapEx_{\F, H, k, \eps}$}

\reporting*

\begin{comment}
\begin{theorem} \label{thm:dvorak-tuma-examples}
	Let $\eps>0$, $H$ be a fixed graph, $\G$ be a graph class of bounded expansion, $k\in \N$, and $\F$ be either $\Hom$, $\Sub$, or $\ISub$. 
	Let $G$ be a dynamic $k$-colored graph on $n$ vertices that belongs to $\G$ at all times.
	Then, there is a randomized data structure $\MapEx_{\F, H, k,\eps}[G]$ which after every update is able to report that either $\F(H, G)$ is empty, or provide some $\phi \in \F(H, G)$. The amortized update time is $\Oh_{\G, H, k}(\log^{\Oh_H(1)} n \log{\frac{1}{\eps}})$, the initialization time is $\Oh_{\G, H, k}(n \log n \log {\tfrac{1}{\eps}})$, and the space complexity is $\Oh_{\G, H, k}(n \log n \log {\tfrac{1}{\eps}})$.  The data structure never provides false positives, but may fail to provide an example mapping $\phi$ with probability at most $\eps$, against an oblivious adversary.
\end{theorem}
\end{comment}

%If $\G$ is a nowhere-dense class, then each update takes $\Oh_{\G, H, k}(|G|^\delta \log{\frac{1}{\eps}})$ time and its space complexity is $\Oh_{\G, H, k}(|G|^{1 + \delta} \log{\frac{1}{\eps}})$.

For the algorithms in the following sections we will only need the case $\F = \Hom$, which is the only case that is easy to handle for the argumentation provided in the work of Dvo\v{r}\'ak and T\r{u}ma~\cite{DvorakT13} (despite not being stated and proven explicitly there). However, the toolbox that we need to introduce for resolving the harder cases of $\F = \Sub$ and $\F = \ISub$ will turn out to be useful later on; and besides, these cases can be also viewed as results of independent interest. In particular, even though the case $\F=\Hom$ does not actually require randomization, in the proofs of \cref{thm:main-domset-be,thm:main-indset-be} we will reuse some randomized parts of the toolbox, and the resulting data structures are randomized.

%\wninline{@Bartek, napiszesz dowód tego lematu? De facto już to zrobiłeś w dom-one i trzeba by go tylko przystosować do języka tutaj. No i przypadek 2 wymaga jeszcze triku z dzieleniem ważonych przez nieważone aby odzyskać indeks}
%\wninline{TODO: Zastanowić się na ile potrzebujemy tak skomplikowanego dowodu}

Let us first introduce a handy definition. A \emph{vertex function} is a function $f$ that takes a graph $G$ and its vertex $v$ and returns a nonnegative integer $f(G,v)$. If we always have $f(G,v)\in \{0,1\}$, we call the vertex function $f$ \emph{binary}. 

The following lemma, which draws inspiration from the works of Majewski, Pilipczuk, and Zych-Pawlewicz~\cite{Majewski24} and of Nadara, Pilipczuk, and Smulewicz~\cite{Nadara22}, is our main tool for recovering examples while given access only to counting. We call this technique \emph{fingerprint retrieval}.
%\wninline{Może skoro kilkukrotnie wspominaliśmy nazwę fingerprint retrieval w overview to ją tutaj wspomnimy aby to ukonkretyzować?}
\begin{lemma}(Fingerprint Retrieval Lemma) \label{lem:restore-by-sampling}
	Let $f$ be a vertex function and let $G$ be a dynamic graph on $n$ vertices.
	Suppose that for any fixed weight function $\wei \colon V(G) \to \Z$, there is a data structure $\WeiSum_{f, \wei}[G]$ that reports the quantity $\sum_{v \in V(G)} f(G, v)\cdot \wei(v)$ after every update to $G$.
	Also, suppose that
	\begin{enumerate}[label={(\arabic*)}]
		\item \label{it:binary-f} $f$ is binary, or
		\item \label{it:general-f} there is a data structure $\Ver_{f}[G]$ that supports the following queries: for a given vertex $v$, decide whether $f(G, v)>0$.
	\end{enumerate}
	Assume further that for the data structure $\WeiSum_{f, \wei}[G]$ and, in case \ref{it:general-f}, also for $\Ver_f[G]$, we have the following complexity guarantees: amortized update and query time at most $T$, initialization time at most~$I$, and space complexity at most $M$.

	Then, for every~$\eps>0$ there exists a randomized data structure $\RetVer_{f,\eps}[G]$ that support the following query: return any $v$ such that $f(G, v) > 0$, or $\bot$ if no such vertex exists. The amortized update and query time is $\Oh(T\log n\log \tfrac{1}{\eps})$, the initialization time is $\Oh((I+n)\log n\log \tfrac{1}{\eps})$, and the space complexity is $\Oh((M+n)\log n\log \tfrac{1}{\eps})$. The data structure never provides false positive, but may fail to provide an example vertex $v$ with probability at most $\eps$, against an oblivious adversary.
\end{lemma}	
	
	
	
	%If the initialization time, amortized time complexity per a single update, amortized query time, and space complexity of $\WeiSum_{f, w}(G)$ are $\Oh(I_R)$, $\Oh(T_R)$, $\Oh(Q_R)$ and $\Oh(S_R)$, respectively, and in the case \cref{it:general-f} the initialization time, amortized time complexity per a single update, amortized query time, and space complexity of $\Ver_f(G)$ are $\Oh(I_V)$, $\Oh(T_V)$, $\Oh(Q_V)$ and $\Oh(S_V)$, respectively, then the initialization time, amortized time complexity per a single update, amortized query time, and space complexity of $\RetVer_f(G)$ are $\Oh((I_R+I_V+|G|) \cdot \log |G| \cdot \log \frac{1}{\eps})$, $\Oh((T_R+T_V) \cdot \log |G| \cdot \log \frac{1}{\eps})$, $\Oh((Q_R+Q_V) \cdot \log |G| \cdot \log \frac{1}{\eps})$ and $\Oh((S_R+S_V+|G|) \cdot \log |G| \cdot \log \frac{1}{\eps})$, respectively, where $\eps$ is the probability of a false negative and where we assume that $I_V = T_V = Q_V = S_V = 0$ in the case \cref{it:binary-f}, where $\Ver$ is not required.

\input{lemma-restore-by-sampling-explicite.tex}

\bigskip

Before proceeding, let us note a useful property of mappings that is shared by homomorphisms, subgraph isomorphisms, and induced subgraph isomorphisms, which essentially says that roots can be gadgeted using edges of unique colors. In the following, by a \emph{type of mappings} we mean a family of functions whose domain is the vertex set of one graph $H$ and the co-domain is the vertex set of another graph $G$.

\begin{definition}
	We say that a type of mappings $\F$ is \textit{root-gadgetable} if it satisfies the following property: Let $\rt{H}{x}$ and $\rt{G}{u}$ be $k$-colored graphs, for some $k\in \N$, each equipped with one root. Let $H'$ be the graph $H$ modified by adding pendants $x'$ and $x''$ adjacent to $x$, and  $G'$ be the graph $G$ modified by adding pendants $u', u''$ adjacent to $u$, so that edges $xx'$ and $uu'$ have color $k+1$ and edges $xx''$ and $uu''$ have color $k+2$. (Thus, $H'$ and $G'$ are $(k+2)$-colored.) Then
	\[\F(H',G') = \F(\rt{H'}{x,x',x''},\rt{G'}{u,u',u''}),\]
	and there is a natural bijection between $\F(H', G')$ and $\F(\rt{H}{x},\rt{G}{u})$ given by restricting any $\phi \in \F(H', G')$ to  $V(H)$. 
\end{definition}

\begin{lemma}\label{lem:root-gadget}
	Homomorphisms, subgraph isomorphisms, and induced subgraph isomorphisms are root-gadgetable types of mappings.
\end{lemma}
\begin{proof}
	Let $\F \in \{\Hom, \Sub, \ISub\}$. Let $\phi \in \F(H', G')$. As (induced) subgraph isomorphisms are also homomorphisms, we have that $\phi$ is a homomorphism from $H$ to $G$. As $x$ and $u$ are the only vertices simultaneously incident to edges of colors $k+1$ and $k+2$ in $H'$ and $G'$, respectively, we must have that $\phi(x)=u$, and consequently also $\phi(x') = u'$ and $\phi(x'') = u''$. It follows that $\F(H',G') = \F(\rt{H'}{x,x',x''},\rt{G}{u,u',u''})$. One can then readily verify that restricting any member of $\F(H', G')$ to $V(H)$ defines a bijection between $\F(H', G')$ and $\F(\rt{H}{x}, \rt{G}{u})$, for any $\F$ from $\{\Hom, \Sub, \ISub\}$.
\end{proof}

We now proceed to the proof of \cref{thm:dvorak-tuma-examples}, which will consist of two steps. First, we are going to design a data structure $\MapVer$ that can identify a possible image of a single vertex $x$ of $H$. Then, we are going to use it repeatedly to identify images of all the vertices of $H$ one by one.

Let us focus on designing $\MapVer$ first. The idea  is to apply \cref{lem:restore-by-sampling} for a vertex function $f$ defined as follows: $f(G, u)$ is the number of mappings from $\F$ that map $x$ to $u$. However, in order to follow this path we have to enrich \cref{thm:dvorak-tuma} with a capability of counting \emph{weighted} mappings. The following definition will be useful.

\begin{definition} \label{def:mapping-value}
	Let $H$ and $G$ be graphs and $\wei\colon V(H)\times V(G)$ be a weight function defined on pairs of vertices: one from $H$ and one from $G$. Let $\phi \colon V(H) \to V(G)$ be a mapping. We define the \textit{value} of $\phi$ (with respect to $\wei$) as follows: \[\val_\wei(\phi) = \prod_{x \in V(H)} \wei(x, \phi(x)).\] Consequently, for a type of mappings $\F$, we define $\val_\wei(\F(H, G))$ as $\sum_{\phi \in \F(H, G)} \val_\wei(\phi)$. This definition can be extended to mappings between rooted graphs in the expected way.
\end{definition}

With this definition, we state a weighted version of \cref{thm:dvorak-tuma} that will help us in applying \cref{lem:restore-by-sampling}. 

\begin{theorem} \label{lem:weighted-dvorak}
	Let $k\in \N$, $H$ be a fixed $k$-colored graph, $\G$ be a graph class of bounded expansion, and $\F$ be either $\Hom$, $\Sub$, or $\ISub$. Let $G$ be a dynamic $k$-colored graph on $n$ vertices that belongs to $\G$ at all times, and let $\wei \colon V(H) \times V(G) \to \Z$ be a fixed weight function. Then there is a data structure $\WeiMaps_{\F, H, \wei}[G]$ that is able to determine $\val_\wei(\F(H, G))$ after each update. The amortized update time is $\Oh_{\G, H, k}(\log^h n)$, where $h = {|H| \choose 2} - 1$, while the initialization time and the space complexity is~$\Oh_{\G, H, k}(n)$. %If $\G$ is a nowhere-dense class, then each update takes $\Oh_{\G, H, k}(|G|^\delta)$ time and its space complexity is $\Oh_{\G, H, k}(|G|^{1 + \delta})$.
\end{theorem}

\cref{lem:weighted-dvorak} can be proved by a rather straightforward modification of the proof of \cref{thm:dvorak-tuma} due to Dvo\v{r}\'ak and T\r{u}ma.
On a high level, \cref{thm:dvorak-tuma} is proven by means of consecutive reductions:
\begin{itemize}
	\item counting induced subgraph isomorphisms is reduced to counting subgraph isomorphisms;
	\item counting subgraph isomorphisms is reduced to counting homomorphisms; and
	\item counting homomorphisms is reduced to counting homomorphisms in a special case, where both $H$ and $G$ are additionally directed, but $H$ is a so-called \textit{elder graph}.
\end{itemize}
The first two reductions use the Inclusion--Exclusion Principle, the third reduction directly expresses general homomorphism counts as sums over homomorphism counts from elder graphs, while homomorphism counts from elder graphs are maintained by maintaining suitable dynamic programming tables.
As can be expected, it is a routine task to incorporate weighted counting into all these steps, in particular into the dynamic programming. For the sake of completeness, in \cref{sec:app-weighted-dvorak} we provide a detailed exposition of the modifications that need to be applied to the proof of \cref{thm:dvorak-tuma} in order to derive  \cref{lem:weighted-dvorak}.

With \cref{lem:weighted-dvorak} in hand, we are now able to design the promised $\MapVer$ data structure:

\begin{lemma} \label{lem:map-vertex}
	Let $\eps>0$, $k\in \N$, $\rt{H}{x}$ be a fixed $k$-colored graph with one root $x$, $\G$ be a graph class of bounded expansion, and $\F$ be either $\Hom$, $\Sub$, or $\ISub$. 
	Let $G$ be a dynamic $k$-colored graph on $n$ vertices that belongs to $\G$ at all times.
	Then, there is a randomized data structure $\MapVer_{\F, \rt{H}{x}, k,\eps}[G]$ which after every update is able to report that $\F(H,G)=\emptyset$, or provide a vertex $u$ of $G$ such that $\F(\rt{H}{x},\rt{G}{u})$ is non-empty. The amortized update time is $\Oh_{\G, \rt{H}{x}, k}(\log^{\Oh_{\rt{H}{x}}(1)} n \log{\frac{1}{\eps}})$, the initialization time is $\Oh_{\G, \rt{H}{x}, k}(n \log n \log {\tfrac{1}{\eps}})$, and the space complexity is $\Oh_{\G, \rt{H}{x}, k}(n \log n \log {\tfrac{1}{\eps}})$. The data structure never provides false positives, but may fail to provide an example vertex $u$ with probability at most $\eps$, against an oblivious adversary.
\end{lemma}
\begin{proof}
	Working towards an application of \cref{lem:restore-by-sampling}, we are going to construct the required data structures $\WeiSum_{f, \wei}[G]$ and $\Ver_f[G]$ such that the resulting data structure $\RetVer_{f,\eps}[G]$ fits the requirements for $\MapVer_{\F, \rt{H}{x}, k,\eps}[G]$.
		
	Let $f$ be a vertex function defined as follows: for a graph $G$ and a vertex $u$ of $G$, 
	\[f(G,u)\coloneqq |\F(\rt{H}{x},\rt{G}{u})|.\] 
	Consider any weight function $\wei\colon V(G)\to \Z$ and define a weight function $\wei' \colon V(H) \times V(G) \to \Z$ in the following way: for $(y,v)\in V(H)\times V(G)$,
	\[
	\wei'(y,v) =
	\begin{cases}
		\wei(v) & \text{if } y = x, \\
		1  & \text{otherwise.}
	\end{cases}
	\]
	Note that for $\wei'$ defined in this way, we have  \[\val_{\wei'}(\F(H, G)) =\sum_{\phi\in \F(H,G)} \wei(\phi(x)) = \sum_{u\in V(G)} |\F(\rt{H}{x},\rt{G}{u})|\cdot \wei(u)= \sum_{u \in V(G)} f(G, u)\cdot \wei(u).\]
	Hence, we can set $\WeiSum_{f, \wei}[G] \coloneqq \WeiMaps_{\F, H, \wei'}[G]$, provided by \cref{lem:weighted-dvorak}.
	
	As for designing $\Ver_f[G]$, let us first construct $H'$ from $H$ by adding two pendants $x',x''$ connected to $x$ by edges $xx'$ and $xx''$ of colors $k+1$ and $k+2$, respectively. Further, for any $u \in V(G)$, let $G_u$ be obtained from $G$ by adding two pendants $u',u''$ connected to $u$ by edges $uu'$ and $uu''$ of colors $k+1$ and $k+2$, respectively. As $\F$ is a root-gadgetable type of mappings by \cref{lem:root-gadget}, we have that $|\F(\rt{H}{x}, \rt{G}{u})|=|\F(H',G_u)|$. Hence, deciding whether $f(G, u)>0$ is equivalent to deciding whether $|\F(H', G_u)| > 0$. Consequently, the data structure $\Ver_f[G]$ can be implemented in the following way. Let us initialize the data structure $\UnMaps_{\F, H'}$, provided by \cref{thm:dvorak-tuma}, on the vertex set $V(G) \cup \{p_1, p_2\}$, where $p_1$ and $p_2$ are fresh vertices that will normally stay isolated. All updates to $G$ are relayed to that data structure. Whenever a query about a vertex $u$ is issued to $\Ver_f(G)$, we temporarily add edges $up_1$ and $up_2$ with colors $k+1$ and $k+2$ to $\UnMaps_{\F, H'}$, in order to transform $G$ to $G_u$, and answer that $f(G, u)>0$ if and only if $|\F(H', G_u)| > 0$. After answering the query, the edges $up_1$ and $up_2$ are removed, thus making $p_1$ and $p_2$ isolated again.
	
	If $\G'$ is the class of all the graphs obtainable by repeatedly adding pendants to a graph from $\G$, then $G_u \in \G'$ and $\G'$ has bounded expansion by \cref{obs:pendants-bnd-exp}. Hence, the amortized time complexity of both handling an update to $G$ or answering a single query by $\Ver_f[G]$ is $\Oh_{\G', H', k+2}(\log^{\Oh_{H'}(1)} |G'|) = \Oh_{\G, \rt{H}{x}, k}(\log^{\Oh_{\rt{H}{x}}(1)} n)$. Similarly, the space complexity and the initialization time is $\Oh_{\G, H, k}(n)$. 
	%Whereas, if $\G$ is nowhere-dense, then the time complexities of handling an update to $G$ or answering a single query are $\Oh_{\G, H, k}(|G|^\delta)$, while its space complexity is $\Oh_{\G, H, k}(|G|^{1 + \delta})$.
	
	We may now
	apply \cref{lem:restore-by-sampling} by supplying it with functions $f, \wei$, data structures $\WeiSum_{f, \wei}[G]$ and $\Ver_f[G]$, and the error parameter $\eps$. This yields the data structure $\RetVer_{f,\eps}[G]$ that may serve as $\MapVer_{\F, \rt{H}{x}, k,\eps}[G]$ with the desired properties.
\end{proof}

%\wninline{Napisały mi się lematy w takiej kolejności w jakiej mi się napisały, ale myślę, że może lepiej jednak je pozamieniać kolejnością, aby nie mieć przydługiego założenia o istnieniu MapVertex z abstrakcyjną złożonością i dopiero potem dowodzić jego istnienia, tylko jednak najpierw udowodnić jego istnienie, a potem już operować na konkretach?}
%\wninline{Zdecydowałem się dodać po dwa pendanty, a nie po jednym, aby nie było problemów z izolowanymi wierzchołkami}
%\begin{lemma} \label{lem:vertices-to-mappings}
%% Then, the data structure is able to either provide a vertex $u \in V(G)$ such that there exists $\phi$ such that $\phi(v)=u$ and $\phi \in \F(H, G)$, or state that such $\phi$ does not exist. Any update to $G$ is handled in amortized $\Oh_{\G, H, k}(\log^{g(|H|, k)}(|G|))$ time complexity (for some monotonic function $g$) and the queries never provide false positives, but may provide false negatives with probability at most $\eps$.
%Let $H$ be a fixed graph, $\G$ be a class of graphs, $G \in \G$ be a dynamic graph and let $\F \in \{\Hom, \Sub, \ISub\}$. Then, there exists a data structure $\MapEx_{\F, H, k}(G)$ that provides $\phi$ such that $\phi \in \F(H, G)$, or state that it does not exist. Any update to $G$ is handled in amortized $\Oh_{\G, H, k}(\log^{g(|H|)}(|G|))$ time complexity for some function $g'$. The queries never provide false negatives, but may provide false negatives with probability at most $|H| \eps$.
%	
%\end{lemma}

We are now ready to conclude the proof of the \cref{thm:dvorak-tuma-examples}.

\begin{proof} [Proof of \cref{thm:dvorak-tuma-examples}]
	Let us enumerate $V(H)$ as $\{x_1, \ldots, x_{|H|}\}$ and create a sequence of graphs $H_0, \ldots, H_{|H|}$, where $H=H_0$ and $H_{i}$ is created from $H_{i-1}$ by adding pendants $p_{2i-1}$ and $p_{2i}$, connected to $v_i$ by edges of colors $k+2i-1$ and $k+2i$, respectively. Let also $\eps'\coloneqq \tfrac{\eps}{|H|}$ and $\G'$ be the class of graphs obtainable from graphs from $\G$ by repeatedly adding pendants. By \cref{obs:pendants-bnd-exp}, $\G'$ has bounded expansion.
	
	We maintain data structures $\MapVer_i \coloneqq \MapVer_{\F, \rt{H_i}{x_{i+1}}, k + 2i,\eps'}$ for $i=0, \ldots, |H|-1$, provided by \cref{lem:map-vertex} for the class $\G'$. Each data structure $\MapVer_i$ is initialized on $n+2i$ vertices: the $n$ vertices of $G$ and $2i$ additional vertices that will normally stay isolated. Any update to $G$ is passed to all of the data structures $\MapVer_0, \ldots, \MapVer_{|H|-1}$, so that each data structure $\MapVer_i$ maintains a graph consisting of $G$ plus $2i$ isolated vertices. 
	
	The data structure $\MapVer_i$ will be used to identify the image of $x_{i+1}$ after already fixing images of $x_1, \ldots, x_i$.
	That is, let us assume that we have already identified vertices $u_1, \ldots, u_i \in V(G)$ such that there exists a mapping $\phi \in \F(H, G)$ with $\phi(x_j)=u_j$ for all $j \le i$. Let $G_0, G_1, \ldots, G_i$ be the sequence of graphs such that $G=G_0$ and $G_i$ is created from $G_{i-1}$ by adding pendants $q_{2i-1}$ and $q_{2i}$ connected to $u_i$ by edges of  colors $k+2i-1$ and $k+2i$, respectively. By using the $2i$ isolated vertices in the graph stored in $\MapVer_i$ and temporarily adding $2i$ edges incident to them, we can turn the graph stored in $\MapVer_i$ into $G_i$. Since $\F$ is root-gadgetable (by \cref{lem:root-gadget}), a straightforward induction on $i$ shows that there is a natural bijection between mappings $\phi' \in \F(H_i, G_i)$ and mappings $\phi \in \F(H, G)$ satisfying $\phi(x_j)=u_j$ for all $j\leq i$, defined by $\phi'|_{V(H)} = \phi$. As we assumed that there exists $\phi \in \F(H, G)$ such that $\phi(x_i) = u_i$, we conclude that $\F(H_i, G_i)$ is nonempty. Hence, $\MapVer_i$ is able to provide a vertex $u_{i+1}$ such that there exists $\phi' \in \F(H_i, G_i)$ with $\phi'(x_{i+1})=u_{i+1}$, or conclude that no such vertex exists. Moreover, if such $u_{i+1}$ is found, then it has to be the case that $u_{i+1} \in V(G)$ (that is, it is not possible that $u_{i+1}=q_j$ for some $j\le i$), so by restricting $\phi'$ to $V(H)$ we get a mapping $\phi$ such that $\phi(x_j)=u_j$ for all $j\le i+1$. By repeating this reasoning for $i=0, \ldots, |H|-1$, we get a full mapping $\phi \in \F(H, G)$, as desired; or a conclusion at some step that no such mapping exists.
	
	Note that we have $|H_i|\leq |H|+2i\leq 3|H|$ and $|G_i| = n+2i\leq n + 2|H|$. Therefore, the promised guarantees on the amortized update time, initialization time, and space complexity follow directly from the guarantees for the data structures $\MapVer_i$, provided by \cref{lem:map-vertex}. Clearly the constructed data structure provides no false positives, and since we make $|H|$ queries to the $\MapVer_i$ data structures in total, the probability of a false negative is bounded by $|H|\eps'=\eps$.
\end{proof}

Finally, we remark that since $|\F(H, G)| \le |G|^{|H|}$ and $\val_\wei(\phi) \le |G|$ for any $\F$, $\wei$, and $\phi$ considered in the proofs presented above, all the integers involved in the computations are of bitlength $\Oh(|H|\log n)$, and hence they fit into $\Oh(|H|)$ words in the RAM model. Therefore, all the arithmetic operations can be performed in time $\Oh_H(1)$.

\subsection{Extension to rooted relational structures}

In the next section we will need a variant of \cref{thm:dvorak-tuma-examples} for rooted, colored directed graphs. While it is easy to add those features to the proof of \cref{thm:dvorak-tuma-examples}, for the sake of providing a robust citation interface for future works we choose to take an extra mile and lift \cref{lem:weighted-dvorak,thm:dvorak-tuma-examples} to the setting of rooted relational structures. The proof is a reduction by means of rather simple gadgeteering, similar to the arguments proposed by Dvo\v{r}\'ak and T\r{u}ma in~\cite[Section~4]{DvorakTumaArxiv}, where they argued that \cref{thm:dvorak-tuma} lifts to this setting.

Let us clarify the model. For a signature $\Sigma$, by a \emph{dynamic $\Sigma$-structure} we mean a $\Sigma$-structure $\Af$ whose universe stays fixed, but which is modified by updates of the following kind: given a relation $R\in \Sigma$ and a tuple $\tup u\in \Af^{\ar(R)}$, add/remove $\tup u$ from $R^\Af$. Data structures for dynamic $\Sigma$-structures are defined analogously to the graph setting.


%There is a natural question to ask about whether \cref{thm:dvorak-tuma-examples} extends to more general settings, e.g. can we ask about directed edges or hyperedges? The short answer is affirmative. In fact, in the following sections we are going to use a generalization to retrieving example homomorphisms between directed graphs. However, due to a technical nature of this argument, we defer precise statement and its proof to \cref{app:relations}. 

\begin{theorem}\label{thm:homcounts-relational}
	Let $\Cc$ be a graph class of bounded expansion, $\Sigma$ be a signature, $\rt{\Bf}{\tup x}$ be a fixed $\Sigma$-structure with roots $\tup x$, and $\F$ be either $\Hom$, $\Sub$, or $\ISub$. Let $\Af$ be a dynamic $\Sigma$-structure with a universe of size $n$ whose Gaifman graph belongs at all times to $\Cc$. Further, let $\wei\colon U(\Bf)\times U(\Af)\to \Z$ be a fixed weight function. Then there is a deterministic data structure $\WeiMaps_{\F, \rt{\Bf}{\tup x},\wei}[\Af]$ for $\Af$ that supports queries:
	\begin{itemize}
		\item Given a tuple $\tup u\in U(\Af)^{|\tup x|}$, compute $\val_\wei(\F(\rt{\Bf}{\tup x},\rt{\Af}{\tup u}))$.
	\end{itemize}
	Moreover, for every $\eps>0$, there is a randomized data structure $\MapEx_{\F, \rt{\Bf}{\tup x},\eps}[\Af]$ for $\Af$ that supports queries:
	\begin{itemize}
		\item Given a tuple $\tup u\in U(\Af)^{|\tup x|}$, return any mapping $\varphi\in\F(\rt{\Bf}{\tup x},\rt{\Af}{\tup u})$, or $\bot$ if this set is empty.
	\end{itemize}
	The amortized update and query time of the data structure $\WeiMaps_{\F, \rt{\Bf}{\tup x},\wei}[\Af]$ is $\Oh_{\Cc}(\log^{\Oh_{\rt{\Bf}{\tup x},\Sigma}(1)} \log n)$, while the initialization time and the space complexity is $\Oh_{\Cc,\rt{\Bf}{\tup x},\Sigma}(n)$. 
	The amortized update and query time of the data structure $\MapEx_{\F, \rt{\Bf}{\tup x},\eps}[\Af]$ is $\Oh_{\Cc}(\log^{\Oh_{\rt{\Bf}{\tup x},\Sigma}(1)} \log n \log \tfrac{1}{\eps})$, while the initialization time and the space complexity is $\Oh_{\Cc,\rt{\Bf}{\tup x},\Sigma}(n\log n\log \tfrac{1}{\eps})$. The data structure $\WeiMaps_{\F, \rt{\Bf}{\tup x},\wei}[\Af]$ always provides correct answers to the queries. The data structure $\MapEx_{\F, \rt{\Bf}{\tup x},\eps}[\Af]$ never outputs false positives, but may output a false negative with probability at most $\eps$ against an oblivious adversary.
\end{theorem}
	
	%Let $H$ and $G$ be two relational structures with $U_H$ and $U_G$ as their universes equipped with relations $R_1, \ldots, R_r$, where the arity of $R_i$ is $c_i$. $H$ is a fixed relational structure, i.e. $R_i(H)$ stays fixed, while $G$ is a dynamic relational structure, i.e. it receives a series of additions and removals to any of $R_1(G), \ldots, R_r(G)$. Let $\F \in \{\Hom, \Sub, \ISub\}$. Let $\eps \in (0, 1)$ be a real number. Then, there exist deterministic data structure $\WeiMaps_{\F, H}(G)$, which is able to determine $|\F(H, G)|$ after each update, and a randomized data structure $\MapEx_{\F, H, \eps}(G)$, which is able to report any $\phi \in \F(H, G)$ if it is nonempty. Let $\G$ be a class of graphs such that $\mathsf{Gaif}(G) \in \G$ at any point in time.
	
	%If $\G$ is of bounded expansion, then the data structure $\mathsf{UnweightedMappings}_{\F, H}(G)$ processes each update in $\Oh_{\G, H}(\log^{g(H)} |U_G|)$ amortized time complexity and requires $\Oh_{\G, H}(|U_G| \log^{g(H)} |U_G|)$ space complexity, while $\MapEx_{\F, H \eps}(G)$ processes each update in $\Oh_{\G, H}(\log^{g(H)} |U_G| \log{\frac{1}{\eps}})$ amortized time complexity and requires $\Oh_{\G, H}(|U_G| \log^{g(H)} |U_G| \log{\frac{1}{\eps}})$ space complexity. 
	
%	If $\G$ is a nowhere-dense class, then the data structure $\mathsf{UnweightedMappings}_{\F, H}(G)$ processes each update in $\Oh_{\G, H}(|U_G|^\delta)$ amortized time complexity and requires $\Oh_{\G, H}(|U_G|^\delta)$ space complexity, while $\MapEx_{\F, H, \eps}(G)$ processes each update in $\Oh_{\G, H}(|U_G|^\delta \log{\frac{1}{\eps}})$ amortized time complexity and requires $\Oh_{\G, H}(|U_G|^\delta \log{\frac{1}{\eps}})$ space complexity. 
	
	%The data structure $\MapEx_{\F, H, \eps}(G)$ provides no false positives, but may fail to provide $\phi$ with probability at most $\eps$.
\begin{proof}
	%We are going to mainly follow the proof of \cite{DvorakTumaArxiv}, but apply a few required changes. We are going to (almost) reduce the problem to the cases of colored undirected graphs already covered by \ref{lem:weighted-dvorak} and \ref{thm:dvorak-tuma-examples}.
	We first argue that we may focus on the case where there are no roots ($\tup x$ and $\tup u$ are empty tuples). Indeed, if $|\tup x|=k$, then we may add $k$ new unary relations $Q_1,\ldots,Q_k$ to the signature and mark the $i$th root $x_i$ of $\rt{\Bf}{\tup x}$ using $Q_i$. Upon query about a tuple $\tup u$ in $\Af$ (either in $\WeiMaps_{\F, \rt{\Bf}{\tup x},\wei}[\Af]$ or in $\MapEx_{\F, \rt{\Bf}{\tup x},\eps}[\Af]$), we may temporarily mark each $u_i$ using $Q_i$, answer the query for unrooted but $\{Q_1,\ldots,Q_k\}$-decorated $\Bf$ and $\Af$, and restore the original state of $\Af$ by unmarking back the elements~$u_i$. Thus, from now on we focus on the case where there are no roots.
	
	By adding a unary relation $R$, such that every element of $\Af$ is a singleton tuple
	in $R$, and every element of $\Bf$ is a singleton tuple in $R$, we may assume that every element of $\Bf$ is in some relation.
	
	We first focus on designing the $\WeiMaps_{\F,\Bf,\wei}[\Af]$ data structure.
	Let $C=\Sigma\uplus 2^{[q]}$, where $q\coloneqq \max_{R\in \Sigma} \ar(R)$.
	Observe, that $q \in \Oh(\Cc)$ is a constant as $\Cc$ has bounded expansion and
	every non-empty $q$-ary relation induces a $q$-clique in the Gaifman graph.
	Let $M_\Bf$ be a $C$-colored undirected graph defined in the following way. First, we include the universe $U(\Bf)$ in the vertex set of $M_{\Bf}$. Then, for each $R\in \Sigma$ and $\tup v\in R^\Bf$, we create two new vertices $t^{R,1}_{\tup v}, t^{R,2}_{\tup v} \in V(M_\Bf)$ and add to $M_\Bf$ the following edges:
	\begin{itemize}
		\item edge $t^{R,1}_{\tup v}t^{R,2}_{\tup v}$ of color $R$; and
		\item for each vertex $w\in U(\Bf)$ appearing in $\tup v$, an edge $wt^{R,1}_{\tup v}$ of color $\{j\in [\ar(R)]\mid w=v_i\}$\footnote{In \cite[Section~4]{DvorakTumaArxiv} all edges of the form $wt^{R,1}_{\tup v}$ are of the same color. However, it is necessary to assign them different colors in order not lose information about the position(s) in the tuple $\tup v$ occupied by the vertex $w$.}.
	\end{itemize}
	We define the graph $M_\Af$ analogously. Note the following (an analogous statement is also proved in~\cite{DvorakTumaArxiv}).
	
	\begin{claim}\label{cl:still-bnd}
		There is a class of bounded expansion $\Cc'$, depending only on $\Cc$ and $\Sigma$, such that
		$M_\Af\in \Cc'$.
	\end{claim}
	\begin{claimproof}
		Let $G$ be the Gaifman graph of $\Af$. Since $G\in \Cc$, $G$ is $d$-degenerate where $d\coloneqq \nabla_0(\Cc)$, hence there exists an ordering $\preceq$ of the vertices of $G$ such that every vertex $u$ of $G$ has at most $d$ neighbors that are smaller in $\preceq$. (Such an ordering can be obtained by repeatedly removing from $G$ a vertex of the smallest degree and ordering the vertices by the reverse order of removal.) For any $R\in \Sigma$ and any tuple $\tup v\in R^{\Af}$, let $\mu(\tup v)$ be the $\preceq$-maximal vertex featured in $\tup v$. Note that for every $u\in V(G)$ we have $|\mu^{-1}(u)|\leq |\Sigma|\cdot (d+1)^q$,
		for every tuple $\tup v\in \mu^{-1}(u)$ is entirely contained in the set consisting $u$ and its $\preceq$-smaller neighbors --- which is of size at most $d+1$.
		Hence, it is easy to verify that the map $\eta$ defined~as
		\begin{align*}
			\eta(u)& \coloneqq \{u\} && \textrm{for }u\in U(\Af);\textrm{ and}\\
			\eta(t^{R,1}_\tup v)\coloneqq \eta(t^{R,2}_\tup v)& \coloneqq \{\mu(\tup v)\} && \textrm{for }\tup v\in R^\Af, R\in \Sigma;
		\end{align*} 
		is a congestion-$(2|\Sigma|\cdot (d+1)^q+1)$ depth-$0$ model of $M_\Af$ in $G$. It follows that $M_\Af\in \Minors^{2|\Sigma|\cdot (d+1)^q+1,0}(\Cc)$, which is a class of bounded expansion by \cref{thm:cong-minors}.
%		\wninline{A nie przypadkiem $(2|\Sigma|\cdot (d+1)^q+1)$-congestion? $\cdot 2$ bo dwa wierzchołki per krotka, $+1$ za samego siebie. Albo w ogóle jeszcze $d^{q-1}$, zamiast $(d+1)^q$, bo wiadomo, że $u$ zawsze jest :)}
%		\azinline{co do pierwszego zdania to tak- poprawiam. co do drugiego $u$ zawsze jest ale może byc wiele razy- zostawiam}
	\end{claimproof}
	
	It is now straightforward to see that subgraph isomorphisms from $M_\Bf$ to $M_\Af$ are in one-to-one correspondence to substructure isomorphisms from $\Bf$ to $\Af$. %We leave the easy verification of the following claim to the reader.
	
	\begin{claim}\label{cl:subgraph-works}
		Every substructure isomorphism $\varphi\in \Sub(\Bf,\Af)$ can be uniquely extended to a subgraph isomorphism $\varphi'\in \Sub(M_\Bf,M_\Af)$. Conversely, every subgraph isomorphism $\varphi'\in \Sub(M_\Bf,M_\Af)$ becomes a substructure isomorphism after restricting the domain to $U(\Bf)$.
	\end{claim}
	
	So if we extend the function $\wei \colon U(\Bf) \times U(\Af) \to \Z$ to a function $\wei' \colon V(M_H) \times V(M_G) 
	\to \Z$ by setting $\wei'(x, u) = \wei(x, u)$ for all $(x, u) \in U(\Bf) \times U(\Af)$, and $\wei'(x, u)=1$ for all other pairs $(x,u)$, then
	\[\val_{\wei'}(\Sub(M_\Bf,M_\Af))=\val_\wei(\Sub(\Bf,\Af)).\]
	Consequently, if we maintain the data structure $\WeiMaps_{\Sub, M_\Bf, \wei}(M_\Af)$ provided by \cref{lem:weighted-dvorak} for the class $\Cc'$ given by \cref{cl:still-bnd}, then this data structure can be directly used to answer queries to $\WeiMaps_{\Sub,\Bf,\wei}[\Af]$. Noting that one update to $\Af$ corresponds to at most $q+1$ updates to $M_\Af$, hence the claimed complexity guarantees about $\WeiMaps_{\Sub,\Bf,\wei}[\Af]$ follow directly from the guarantees provided by \cref{lem:weighted-dvorak}. We note that each update to $\Af$, apart from altering the edge set of $M_\Af$, also alters the its vertex set by either adding or removing two vertices specific to the toggled tuple, so it seems as if a method for creating or removing isolated vertices should be required. However, as argued in the \cref{cl:still-bnd}, the number of tuples in $\Af$ can be upper bounded by $n|\Sigma|\cdot (d+1)^q$, so we can deal with that by creating a stash of $2n|\Sigma|\cdot (d+1)^q$ isolated vertices during the initialization. Whenever we insert a tuple $\tup v$ to $R^\Af$, we take two vertices from the stash and name them $t^{R,1}_\tup v$ and $t^{R,2}_\tup v$ and whenever we remove $\tup v \in R^\Af$, we return these to the stash.
	\wninline{Dopisałem tu wyżej kawałek o tych izolowanych}
	
	% then the $w-$weight of the subrelation isomorphism will be the same as the $w'$-weight of the corresponding subgraph isomorphism.
	%As each update to either of $R_i(G)$ incurs a constant number of updates to $M_G$, we easily get the existence of $\WeiMaps_{\Sub, H, w}(G)$ and $\MapEx_{\Sub, H, \eps}(G)$. Also, let $\G'$ be the class of graphs that can be obtained as $M_G$ for $G \in \G$. In \cite{DvorakTumaArxiv} it is proven that if $\G$ is of bounded expansion then $\G'$ is of bounded expansions as well, and if $\G$ is nowhere-dense, then $\G'$ is nowhere-dense as well.
	
	%	\wninline{Pogadac o Gaifmanach, tzn. udowodnić, że $Gaif(G) \in \G$, gdzie $\G$ jest bnd exp/nowhere-dense daje że $M_G \in \G'$, gdzie $\G'$ też jest bnd exp/nowhere-dense}
	
	While \cref{cl:subgraph-works} holds for subgraph isomorphisms, it unfortunately fails for induced subgraph isomorphisms and homomorphisms. Fortunately, it is well-known that weighted homomorphism counts and weighted induced subgraph isomorphism counts can be expressed as linear combinations of weighted subgraph isomorphism counts. Let us explain this argument in more detail, using the argumentation from \cite{DvorakTumaArxiv}, which is repeated in \cref{sec:app-weighted-dvorak}. In the following we assume that the reader is familiar with the material from this section.
	
	%In order to extend this conclusion to $\Hom$ and $\ISub$ we are going to use similar relations between them and $\Sub$ as in the case of graphs, however we are going to apply them directly on the relations rather than on $M_H$ and $M_G$\footnote{We note that in \cite{DvorakTumaArxiv} the authors worked on induced subgraph isomorphisms between corresponding graphs, but this does not lead to a correct algorithm as induced subrelation isomorphisms do not correspond to induced subgraph isomorphisms from $M_H$ to $M_G$.}.
	
	Let us tackle the case of homomorphisms first. Similarly as in \cref{def:projections} for any partition $\Pp$ of $U(\Bf)$, we may define the quotient $\Sigma$-structure $\Bf_\Pp$ by identifying every part $P\in\Pp$ into a single element $z_P$, and pushing the relations in $\Bf$ naturally to relations in $\Bf_\Pp$ through this identification mapping. We may also naturally define the quotient weight function $\wei_\Pp\colon U(\Bf)\times U(\Af)$ by setting $\wei_\Pp(z_P,u)\coloneqq \prod_{x\in P} \wei(x,u)$, for each $P\in \Pp$ and $u\in U(\Af)$. It then follows that 
	\[\val_\wei(\Hom(\Bf, \Af)) = \sum_{\Pp\colon \textrm{partition of }U(\Bf)} \val_{\wei_{\Pp}}(\Sub(\Bf_\Pp, \Af)).\]
	Hence, maintaining the data structures $\WeiMaps_{\Sub, \Bf_\Pp, \wei_\Pp}(\Af)$ for all partitions $\Pp$ pf $U(\Bf)$ allows us to answer the queries to $\WeiMaps_{\Hom, \Bf, \wei}[\Af]$ within the claimed complexity bounds.
	
	For the case of induced substructure isomorphisms we use the same Inclusion--Exclusion argument as in \cref{lem:weighted-isub-to-sub}, but instead of iterating over all supergraphs, it suffices to iterate over all superstructures, that is, all $\Sigma$-structures $\Bf'$ such that $U(\Bf') = U(\Bf)$ and $R^\Bf \subseteq R^{\Bf'}$ for all $R\in \Sigma$. Specifically, we have
	\[\val_\wei(\ISub(H, G)) = \sum_{\Bf'\colon \textrm{superstructure of } \Bf} (-1)^{\|\Bf'\| - \|\Bf\|}\cdot \val_\wei(\Sub(\Bf', \Af)),\]  where $\|\Bf\| = \sum_{R\in \Sigma} |R^\Bf|$, and similarly for $\Bf'$. Since all such superstructures $\Bf'$ are still of size $\Oh_{\Bf,\Sigma}(1)$ and there is $\Oh_{\Bf,\Sigma}(1)$ of them, maintaining the data structures $\WeiMaps_{\Sub, \Bf', \wei}(\Af)$ allows us to answer the queries to $\WeiMaps_{\ISub, \Bf, \wei}[\Af]$ within the claimed complexity bounds.
	
	Having designed a suitable data structure $\WeiMaps_{\F, \Bf, \wei}[\Af]$ for each $\F\in \{\Hom,\Sub,\ISub\}$, we may repeat the argumentation of \cref{lem:restore-by-sampling,lem:map-vertex,thm:dvorak-tuma-examples} to design also the data structure $\MapEx_{\F, \Bf,\eps}[\Af]$. The reasoning lifts essentially verbatim; we leave the details to the reader.
\end{proof}
	
	
	%All graphs obtained from this reduction will still be of constant sizes and there will be a constant number of them, hence the data structures $\WeiMaps_{\ISub, H, w}(G)$ and $\MapEx_{\ISub, H}(G)$ can be designed by reducing them to a constant number of $\WeiMaps_{\Sub, H', w}(G)$ and $\MapEx_{\Sub, H'}(G)$ structures.
	
	%where $\val_{w_{H'}}$ is defined similarly as in \cref{lem:unweighted-sub-to-homo}. Hence, we can design $\WeiMaps_{\Hom, H, w}(G)$ and $\MapEx_{\Hom, H}(G)$ by reducing them to a constant number of corresponding data structures $\WeiMaps_{\Sub, H', w_{H'}}(G)$ and $\MapEx_{\Sub, H'}(G)$.
	
	



%\wninline{Dopisać możliwości rozszerzenia o struktury relacyjne}
%We also remark that \cref{thm:dvorak-tuma-examples} can be easily extended to the setting of considering mappings between relational structures provided that the class of Gaifman graphs of corresponding relational structures is still either of bounded expansion or nowhere-dense. The argument is described the Section 4 of \cite{DvorakTumaArxiv} and it carries seamlessly to our statement as Dvo\v{r}\'ak and T\r{u}ma show that any 
