\section{Progressive exploration}\label{sec:ladders}

In this section we recall the framework of \emph{progressive exploration}, proposed by Fabia\'nski et al.~\cite{FabianskiPST19,Fabianski2018arxiv} for  designing parameterized algorithms for \drds{r} and \dris{r}, among other problems of similar kind. While \cite{FabianskiPST19} is the conference version, we refer to the arxiv version \cite{Fabianski2018arxiv} for all the proofs. Our main focus here is to explain that the progressive exploration algorithms for \drds{r} and \dris{r} can be executed using a bounded number of calls to the following two basic queries, each working on a graph $G$ and stipulated by a distance parameter $r\in \N$:
\begin{itemize}
	\item $\far{S,r}$: Given a set $S \subseteq V(G)$, return a vertex $v$ of $G$ such that $\dist_G(v,S)>r$; or $\bot$ if no such vertex exists.
	\item $\near{S,r}$: Given a set $S \subseteq V(G)$, return a vertex $v$ of $G$ such that $\dist_G(v,s)\leq r$ for all $s\in S$; or $\bot$ if no such vertex exists.
\end{itemize}
For the \dris{r} problem, we will need a slight generalization of these.
This will allow us to concentrate in subsequent sections on implementing those queries efficiently in the setting of dynamic data structures.

\subsection{Domination}
For \drds{r}, Fabia\'nski et al. proposed a \emph{semi-ladder algorithm} that proceeds as follows (see \cref{alg:1} for a pseudo-code). The algorithm is employed on a graph $G$ and searches for a distance-$r$ dominating set of size $k$, for given parameters $r,k\in \N$. It assumes access to two queries:
\begin{itemize}
	\item $\notDominated{D,r}$:
	Given a set $D$ of at most $k$ vertices, return a vertex $v$ that is not distance-$r$ dominated by~$D$, that is, such that $\dist_G(v,D)>r$. In case no such vertex $v$ exists, return $\bot$.
%	\wninline{Nie lubię nazwy $\mathrm{witnessDom}$, bo sugeruje jakby to był świadek \emph{dominacji}, a to świadek \emph{niedominacji}. Podobna uwaga co do $\mathrm{witnessInd}$}
%	\wninline{W sporą konfuzję mnie wprowadziła ta narracja i wording tutaj, że zakładamy istnienie ,,more basic query farVertex'' i zarysowywujemy, że będzie chcieli za jego pomocą jakoś zaimplementować zapewne bardziej skomplikowane ,,witnessDom'', które jest dokładnie tym samym. Rozumiem, że obie nazwy mają sens w swoich kontekstach, ale uważnemu czytelnikowi się zapala lampka, że pewnie czegoś nie zrozumiał.}
	\item $\candidate{W,r,k}$: Given a vertex subset $W$, return a set $D$ of at most $k$ vertices that distance-$r$ dominates~$W$, that is, $\dist_G(w,D)\leq r$ for all $w\in W$. In case no such set $D$ exists, return $\bot$.
\end{itemize}
Note that the $\notDominated{D,r}$ query is just equivalent to the $\far{D,r}$ query. We later~show that also the $\candidate{W,r,k}$ query can be implemented using a bounded (in terms of $k$ and~$|W|$) number of $\near{S,r}$ queries with $S\subseteq W$.

The algorithm iteratively constructs two sequences:
\begin{itemize}
	\item a sequence of \emph{candidates} $D_1,D_2,\ldots$, each being a set of at most $k$ vertices; and 
	\item a sequence of \emph{witnesses} $w_1,w_2,\ldots$, each being a single vertex of $G$.
\end{itemize}
After the $(i-1)$st iteration, candidates $D_1,\ldots,D_{i-1}$ and witnesses $w_1,\ldots,w_{i-1}$ are already constructed. Then the $i$th iteration proceeds as follows:
\begin{itemize}
	\item First, we call $\candidate{W,r,k}$ where $W$ comprises of all the  witnesses gathered so far, that is, $W\coloneqq \{w_1,\ldots,w_{i-1}\}$. This either yields the next candidate $D_i$ that distance-$r$ dominates $W$, or a conclusion that no such $D_i$ exists. In the latter case, we may conclude that $G$ has no distance-$r$ dominating set of size at most $k$.
	\item Second, we call $\notDominated{D_i,r}$ to verify whether $D_i$ already distance-$r$ dominates the whole graph. If so, then we have found a solution, and otherwise the call provides the next witness $w_i$, with which we may proceed to the next iteration.
\end{itemize}
It is clear that when this algorithm returns an answer, then this answer is always correct. The main insight of Fabia\'nski et al. is that since every candidate $D_i$ distance-$r$ dominates all the witnesses $w_1,\ldots,w_{i-1}$ but not the witness $w_i$, the constructed candidates and witnesses form a pattern called a \emph{semi-ladder}. Since semi-ladders cannot be too long in classes of bounded expansion, this provides a bound on the number of iterations that the algorithm executes before breaking the loop and reporting an outcome.


\begin{algorithm}
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	
	\nonl
	\underline{procedure DomSet} $(G,r,k)$\\
	\Input{A graph $G$ and parameters $r,k\in \N$}
	\Output{A distance-$r$ dominating set of size $k$, or $\bot$ if no such set exists.}
	\medskip
	$W \coloneqq \emptyset$\;
	\Repeat
	{
		$D\coloneqq \candidate{W,r,k}$\;
		\If{$D = \bot$}
		{
			\KwRet{$\bot$} \;
		}
		$w \coloneqq \notDominated{D,r}$ \;
		\If{$w = \bot$}
		{
			\KwRet{$D$} \;
		}
		$W \coloneqq W \cup \{ w \}$ \;
	}
	\caption{Semi-ladder algorithm for distance-$r$ dominating set of size $k$ in a graph $G$}\label{alg:1}
\end{algorithm}

To be more precise, a \emph{distance-$r$ semi-ladder} of order $\ell$ in a graph $G$ is a pair of sequences of vertices $a_1,\ldots,a_\ell$ and $b_1,\ldots,b_\ell$ such that
\begin{itemize}
	\item $\dist_G(a_i,b_i)>r$ for each $i\in [\ell]$; and
	\item $\dist_G(a_i,b_j)\leq r$ for all $1\leq j<i\leq \ell$.
\end{itemize}
The \emph{distance-$r$ semi-ladder index} of a graph $G$, denoted $\sli_r(G)$, is the largest order of a distance-$r$ semi-ladder that can be found in $G$. For a graph class $\Cc$, we define its distance-$r$ semi-ladder index as $\sli_r(\Cc)\coloneqq \sup_{G\in \Cc} \sli_r(G)$. Note that this value may be infinite if graphs from $\Cc$ contain distance-$r$ semi-ladders of arbitrarily large order. But as proved by Fabia\'nski et al., this does not happen in bounded expansion~classes.

\begin{theorem}[{\cite[Lemma~29]{Fabianski2018arxiv}}]\label{thm:sli-be}
	For every class $\Cc$ of bounded expansion and $r\in \N$, $\sli_r(\Cc)$ is finite.
\end{theorem}

We remark that \cref{thm:sli-be} holds even in a larger generality when $\Cc$ is nowhere dense, but we will not use this here. We also note that for $r=1$, excluding a biclique suffices to bound the semi-ladder index.

\begin{theorem}[{\cite[Lemma~33]{Fabianski2018arxiv}}]\label{thm:sli-biclique}
	For every graph $G$ that does not contain the biclique $K_{t,t}$ as a subgraph, for some $t\in \N$, we have $\sli_1(G)<3t$.
\end{theorem}

The following statement summarizes how a bound on the semi-ladder index influences the number of iterations executed by the semi-ladder algorithm.

\begin{theorem}[consequence of {\cite[Lemma~5 and Corollary~13]{Fabianski2018arxiv}}]\label{thm:termination}
	Let $\Cc$ be a graph class and $r\in \N$ be such that $\sli_r(\Cc)$ is finite. Then the semi-ladder algorithm (\cref{alg:1}) run on any graph from $\Cc$ terminates after executing less than $k^{\sli_r(\Cc)}$ iterations.
\end{theorem}
\begin{proof}[Proof sketch]
Let us briefly sketch how this result follows from the discussion in~\cite{Fabianski2018arxiv}.

In~\cite{Fabianski2018arxiv}, Fabia\'nski et al. formulate the semi-ladder index of first-order formulas as follows. For a graph $G$ and a first-order formula $\phi(\tup x,\tup y)$, where $\tup x,\tup y$ are tuples of variables, they define $\phi(G)$ be a bipartite graph $(A \cup B, F)$ where $L=V(G)^{\tup x}$ (evaluations of variables of $\tup x$ in $V(G)$, $R=V(G)^{\tup y}$ (same for $\tup y$), and for $\tup a\in V(G)^{\tup x}$ and $\tup b\in V(G)^{\tup y}$, we put $\tup a\tup b \in F$ iff $\phi(\tup a,\tup b)$ holds in $G$.
A \emph{semi-ladder} of order $\ell$ in $\varphi(G)$ is a pair of sequences $\tup a_1,\ldots,\tup a_\ell\in A$ and $\tup b_1,\ldots,\tup b_\ell\in B$ such that $a_ib_i\notin F$ for all $i\in [\ell]$ and $a_ib_j\in F$ for all $i,j\in [\ell]$ with $i<j$. The \emph{semi-ladder index} of $\phi(G)$ is the largest order of a semi-ladder in $\phi(G)$, and the \emph{semi-ladder index} of a formula $\phi(\tup x,\tup y)$ on a graph class $\Cc$ is defined as the supremum of the semi-ladder indices of graphs $\phi(G)$, for $G\in\Cc$.

In this notation, the quantity $\sli_r(\Cc)$ is equal to the semi-ladder index on $\Cc$ of the formula $\delta_r(x,y)$ that verifies whether $x$ and $y$ are at distance at most $r$. \cite[Lemma~5]{Fabianski2018arxiv} then says that the semi-ladder index of the formula $\delta^k_r(\tup x,y)=\bigvee_{i=1}^k \delta_r(x_i,y)$, which checks whether the $k$-tuple of vertices $\tup x$ distance-$r$ dominates $y$, has semi-ladder index smaller than $k^\ell$, where $\ell$ is the semi-ladder index of $\delta_r(x,y)$. With this, \cite[Corollary~13]{Fabianski2018arxiv} asserts that $k^\ell-1$ is an upper bound on the number of iterations executed by the semi-ladder algorithm.
\end{proof}

%For our purposes, the formula $\phi(u,v)$ considered in \cite[Lemma~5]{Fabianski2018arxiv} is logically equivalent to $u$ being at distance at most $r$ from $v$. A distance-$r$ semi-ladder in $G$ precisely corresponds to semi-ladder in $\varphi(G)$. The formula $\psi$ extends $\varphi$ to accept
%$k$-touples as the first argument, in a way that $\psi(\bar{u},v)$ is logically equivalent to $v$ being at distance at most $r$ in $G$ from some vertex of $\bar{u}$, i.e., $v$ being $r$-dominated by $\bar{u}$. The precise statement  of Lemma 5 is that if the semi-ladder index of $\varphi(G)$ (the maximum size of a semi-ladder contained in $\varphi(G)$) is at most $l$, then semi-ladder index of $\psi(G)$ is at most $k^{l-1}$. Then Lemma~24 claims that this is also the number of rounds of the semi-ladder algorithm.

Note here that if the algorithm performs at most $\ell\coloneqq k^{\sli_r(\Cc)}$ iterations in total, then all the calls to the $\candidate{W,r,k}$ query are applied only to sets $W$ of size at most $\ell$.

Our goal now is to reduce implementation of the $\notDominated{D,r}$ and $\candidate{W,r,k}$ queries to the more basic queries $\far{S,r}$ and $\near{S,r}$. For $\notDominated{D,r}$ this is trivial: queries $\notDominated{D,r}$ and $\far{D,r}$ are just equivalent. For $\candidate{W,r,k}$, this is a bit more complex.

\begin{lemma}\label{lem:candidate-to-near}
	Let $G$ be a graph, $W$ be a subset of vertices of $G$, and $k,r\in \N$ be parameters. Then the query $\candidate{W,r,k}$ can be answered in time $\Oh(k^{|W|+1})$ by performing at most $k^{|W|+1}$ queries of the form $\near{S,r}$, where $S$ is a subset of $W$.
\end{lemma}
\begin{proof}
	We enumerate all
	the partitions of $W$ into at most $k$ subsets; there are at most $k^{|W|}$ such partitions. For every partition $\cal P$ and every part $S\in \cal P$, we issue the query $\near{S,r}$ to verify whether there exists a vertex $v_S$ such that $\dist_G(v_S,s)\leq r$ for all $s\in S$. If for some partition $\cal P$ we can find such a vertex $v_S$ for every part $S\in \cal P$, then the set $D\coloneqq \{v_S\colon S\in \cal P\}$ is a valid answer to the query $\candidate{W,r,k}$. And if this check fails for every considered partition $\cal P$, then there is no set $D$ of size at most $k$ that distance-$r$ dominates $W$ and we can safely return $\bot$.
	Thus, the total
	number of queries $\near{S,r}$ issued is bounded by
	$k^{|W|+1}$.
\end{proof}

We conclude this section by a statement summarizing the discussion and implying the following: designing an efficient data structure for the $\near{S,r}$ and $\far{S,r}$ queries suffices to implement the semi-ladder algorithm in the dynamic setting, thereby providing a dynamic data structure for the \drds{r} problem.

\begin{lemma}\label{lem:summary-dom}
	Let $\Cc$ be a class of graphs, $k,r\in \N$ be such that $\sli_r(\Cc)$ is finite, and $\ell\coloneqq k^{\sli_r(\Cc)}$. Let $G$ be a dynamic graph that belongs to $\Cc$ at all times. Suppose that there are data structures $\nearVertexDS{G}{r,\ell}$, supporting queries $\near{S,r}$ in $G$ with $|S|\leq \ell$, and $\farVertexDS{G}{r,k}$, supporting queries $\far{S,r}$ in $G$ with $|S|\leq k$, so that $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ have amortized update/query time $T$, initialization time $I$, and space complexity $M$. Then there is a data structure $\dominatingSetDS{G}{r,k}$ for $G$ that supports the~query
	$\ds{}$, which returns a distance-$r$ dominating set of size at most $k$ in $G$, or $\bot$ if no such set exists.
	The amortized update time of $\dominatingSetDS{G}{r,k}$ is $2^{k^{\Oh(\sli_r(\Cc))}}\cdot T$, the query time is $\Oh(k)$, the initialization time is $\Oh(I)$, and the space complexity is $\Oh(M)$.
	
	Moreover, we may allow that the data structures $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ are randomized with error probability at most $\eps$, for any fixed parameter $\eps>0$ and against an oblivious adversary, with amortized update/query time, initialization time, and space complexity becoming $T\cdot \log \tfrac{1}{\eps}$, $I\cdot \log \tfrac{1}{\eps}$, and $M\cdot \log \tfrac{1}{\eps}$, respectively. In this case,  $\dominatingSetDS{G}{r,k}$ is also randomized with error probability at most $\eps$ against an oblivious adversary. The amortized update time becomes $2^{k^{\Oh(\sli_r(\Cc))}}\cdot T\cdot \log \tfrac{1}{\eps}$, the query time remains $\Oh(k)$, the initialization time becomes $k^{\Oh(\sli_r(\Cc))}\cdot I\cdot \log \tfrac{1}{\eps}$, and the space complexity becomes  $k^{\Oh(\sli_r(\Cc))}\cdot M\cdot \log \tfrac{1}{\eps}$.
\end{lemma}
\begin{proof}
	For simplicity, we first assume that data structures $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ are deterministic. We will then explain how to strengthen the proof if they are randomized.
	
	The data structure $\dominatingSetDS{G}{r,k}$ simply maintains the assumed data structures $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$. Upon every update to $G$, we relay the update to $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ and we run the semi-ladder algorithm on $G$ to find a distance-$r$ dominating set of size at most $k$ (or detect a lack thereof). By \cref{thm:termination}, this algorithm can be implemented using at most $\ell$ queries $\notDominated{D,r}$ with $|D|\leq k$ and at most $\ell$ queries $\candidate{W,r,k}$ with $|W|\leq \ell$. Every query $\notDominated{D,r}$ can be answered by a single call to the query $\far{D,r}$ offered by $\farVertexDS{G}{r,k}$. By \cref{lem:candidate-to-near}, every query $\candidate{W,r,k}$ can be answered by performing $k^{|W|+1}\leq 2^{k^{\Oh(\sli_r(\Cc))}}$ calls to queries of the form $\near{S,r}$ with $S\subseteq W$, offered by $\nearVertexDS{G}{r,\ell}$. Once a distance-$r$ dominating set of size $k$ (or lack thereof) is computed, it can be provided in time $\Oh(k)$ upon any $\ds{}$ query.
	The claimed bounds on the amortized update time of $\dominatingSetDS{G}{r,k}$ follow immediately from the assumed guarantees about $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$; and similarly for the initialization time and the space complexity.
	
	In case $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ are randomized, we first assume that they work against an adaptive adversary and we will then explain how to improve the argument if they only work against an oblivious adversary. The data structure $\dominatingSetDS{G}{r,k}$ maintains their instances with a rescaled error parameter $\eps'\coloneqq \tfrac{\eps}{q}$, where
	\[q\coloneqq k^{\sli_r(\Cc)}+k^{k^{\sli_r(\Cc)}}\leq 2^{k^{\Oh(\sli_r(\Cc))}}\]
	is an upper bound on the total number of queries to $\nearVertexDS{G}{r,\ell}$ or $\farVertexDS{G}{r,k}$ that may occur in a single run of the semi-ladder algorithm. Therefore, by the union bound, we conclude that with probability at least $1-\eps$ none of these queries returns an incorrect answer and the semi-ladder algorithm runs correctly. It is straightforward to verify that rescaling the error probability in $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ influences the complexity guarantees for $\dominatingSetDS{G}{r,k}$ as claimed.
	

	We remark that even if the adversary issuing the updates to our graph is oblivious, this argument still required $\nearVertexDS{G}{r,\ell}$ and $\farVertexDS{G}{r,k}$ to work against an adaptive adversary, because the semi-ladder algorithm acts as an internal adaptive adversary --- it issues a number of $\mathtt{nearVertex}/\mathtt{farVertex}$ queries that depend on the answers to previous $\mathtt{nearVertex}/\mathtt{farVertex}$ queries in that execution of the \cref{alg:1}. To remedy this, we initialize multiple independent instances of our helper data structures: $\nearVertexDSS{G}{r,\ell}{1}, \ldots, \nearVertexDSS{G}{r,\ell}{\ell}, \farVertexDSS{G}{r,k}{1}, \ldots, \farVertexDSS{G}{r,k}{\ell}$. In the $i$th iteration of the main loop of the \cref{alg:1} we use  $\nearVertexDSS{G}{r,\ell}{i}$ and $\farVertexDSS{G}{r,k}{i}$ data structures to answer any $\mathtt{nearVertex}$ and $\mathtt{farVertex}$ queries asked during that iteration. The $\mathtt{nearVertex}$ queries asked within the $i$th iteration of different runs do not depend on answers to each other, hence they can all be handled by a single instance $\nearVertexDSS{G}{r,\ell}{i}$ working only against an oblivious adversary; and similarly for the $\farVertexDSS{G}{r,k}{i}$ data structure. This guarantees the correctness in the setting with the weaker assumptions, but it incurs an additional multiplicative $\ell$ factor to update/query time, initialization time, and memory. However, this additional $\ell$ factor does not change the complexities as stated.
\end{proof}

%The algorithm to compute $r$-domination set of size $k$ is then implemented using these procedures
%and the technique called by the authors \emph{progressive exploration}.
%The progressive exploration technique constructs a sequence of candidate solutions. For each
%candidate solution that it constructs, it then tries to invalidate it by finding a witness vertex
%that is not dominated by it. Then the next candidate solution is constructed to dominate both the
%current witness and previously explored witnesses. This process is shown in
%Algorithm~\ref{alg:1}.


%To obtain a bound on the running time of this simple algorithm the
%authors introduce a number of tools from logic and combinatorial graph theory.
%In particular, they define a bipartite graph induced by what they call
%a distance formula. This fomula is a logical sentence stating
%some specific property of a graph $G$ that is an input instance to the
%$r$-dominating set problrem.
%For the $r$-dominating set problem, the
%distance formula is
%$\delta^k_r(\overline{x},y)=\bigvee_{i=1}^k \delta_r(x_i,y)$,
%stating that there is some $x_i$ in a touple
%$\overline{x}=(x_1,\ldots,x_k)$ such that the distance from $x_i$ to $y$
%is at most $r$. This is a First Order Logic formula, whose length
%depends linearly on two parameters: $k$-the size of $\overline{x}$ and
%$r$ - the distance. The bipartite graph $B(\delta^k_r,G)$ induced
%by formula
%$\delta^k_r$ on a graph $G$ is then defined as follows. The left side
%$L$ consists of all tuples $(v_1,\ldots,v_k) \subseteq V^k$ of
%vertices of $G$. The left side $L$ represents
%the universum of all candidate solutions for the $r$-dominating
%set that we seek. The right side $R$ consists of
%all vertices $v \in V(G)$. The right side is viewed as a set of
%witnesses. There is an edge between a candidate
%$(v_1,\ldots,v_k) \in L$ and a witness $v \in R$ if and only if
%$\delta^k_r((v_1,\ldots,v_k),v)$ holds, i.e., $v$ is at
%distance $r$ from some vertex in $(v_1,\ldots,v_k)$. Clearly,
%a feasible solution is a neighbour in $B(\delta^k_r,G)$ to all
%vertices $v \in R$. Unfortunately we cannot afford to construct
%the entire graph $B(\delta^k_r,G)$.
%Algorithm~\ref{alg:1} searches $B(\delta^k_r,G)$ by successively
%exploring new candidates and witnesses, one candidate and one
%witness in each round of the while loop.
%It is easy to observe (see~\cite{FabianskiPST19}), that the
%candidates and witnesses explored by the algorithm induce a
%semi-ladder in $B(\delta^k_r,G)$. Thus, the authors define the
%semi-ladder index $\sli(B)$ of a bipartite graph $B$ as the size of the
%largest induced semi-ladder contained in $B$. Thus, the number of
%rounds executed by Algorithm~\ref{alg:1} is bounded by
%$\sli(B(\delta^k_r,G))$. This in turn, as we shall see later,
%is nicely bounded from above
%for some specific graph classes, in particular bounded expansion graphs.

%By Theorem 11 and Corollary 10 in \cite{FabianskiPST19}, we get that on nowhere dense graphs
%the progressive exploration algorithm (Algorithm~\ref{alg:1})
%performs a bounded number rounds, since the number of rounds corresponds to the size
%of a semi-ladder in a bipartite graph induced by a distance formula
%(see \cite{FabianskiPST19} for the details).

\subsection{Independence}

For the \dris{r} problem, Fabia\'nski et al.~\cite{FabianskiPST19,Fabianski2018arxiv} proposed a different procedure, called the \emph{ladder algorithm}. Similarly to the semi-ladder algorithm, it is based on alternately finding candidates for the sought distance-$r$ independent set and witnesses that the current candidate is not yet a solution. The notion of witnessing is, however, more delicate. For the further discussion, we fix the distance parameter $r\in \N$.

\begin{definition}
	Let $G$ be a graph  and $\tup a$ be a tuple of vertices $G$. We say that a vertex $w\in V(G)$ is a \emph{distance-$r$ dependence witness} for $\tup a$ if there are distinct $i,j\in [|\tup a|]$ such that $\dist_G(a_i,w)+\dist_G(a_j,w)\leq r$. More generally, a set $P\subseteq V(G)$ is a distance-$r$ dependence witness for $\tup a$ if $P$ contains some $w\in P$ that is a distance-$r$ dependence witness for $\tup a$.
\end{definition}

Note that a $k$-tuple $\tup a$ of vertices in a graph $G$ forms a distance-$r$ independent set of size $k$ if and only if there is no distance-$r$ dependence witness for $\tup a$.

The ladder algorithm will be parameterized by one more parameter $p\in \N$, which stipulates the size of dependence witnesses investigated by the algorithm. Finding candidates and witnesses is delegated to the following two queries, which will eventually be implemented using (generalizations of) $\near{S,r}$ and $\far{S,r}$.
\begin{itemize}
	\item $\witnessInd{\cal L,r,p}$:
	Given a family $\cal L$ of tuples of vertices, return a set $P\subseteq V(G)$ of size at most~$p$ such that $P$ is a distance-$r$ dependence witness for all $\tup a\in \cal L$. In case no such vertex set $P$ exists, return $\bot$.
	\item $\candidateInd{W,r,k}$: Given a vertex subset $W$, return $k$-tuple $\tup a$ of vertices such that $W$ is not a distance-$r$ dependence witness for $\tup a$. In case no such tuple $\tup a$ exists, return $\bot$.
\end{itemize}

\begin{algorithm}
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	
	\nonl
	\underline{procedure IndSet} $(G,r,k,p)$\\
	\Input{A graph $G$ and parameters $r,k,p\in \N$}
	\Output{Conclusion whether there exists a distance-$r$ independent set of size $k$ in $G$}
	\medskip
	${\cal L}\coloneqq \emptyset$\;
	$W \coloneqq \emptyset$\;
	\Repeat
	{
		$\tup a\coloneqq \candidateInd{W}$\;
		\If{$\tup a = \bot$}
		{
			\KwRet{``There is no distance-$r$ independent set of size $k$''} \;
		}
		${\cal L}\coloneqq {\cal L}\cup \{\tup a\}$ \;
		$P \coloneqq \witnessInd{\cal L,p}$ \;
		\If{$P = \bot$}
		{
			\KwRet{``There exists a distance-$r$ independent set of size $k$''} \;
		}
		$W \coloneqq W \cup P$ \;
	}
	\caption{Ladder algorithm for distance-$r$ independent set of size $k$ in a graph $G$}\label{alg:2}
\end{algorithm}

%\wninline{Dlaczego ten algorytm jest konsystentnie nazywany ladder algorithm, a domset konsystentnie semi-ladder? Oba robią semi-ladder, tak?}

The ladder algorithm for parameters $r,k,p$ proceeds iteratively in rounds as follows (see \cref{alg:2} for a pseudocode). We construct two sequences: candidates $\tup a_1,\tup a_2,\tup a_3,\ldots$, each being a $k$-tuple of vertices, and witnesses $P_1,P_2,P_3,\ldots$, each being a vertex subset of size at most $p$. After the $(i-1)$st iteration, the candidates $\tup a_1,\ldots,\tup a_{i-1}$ and the witnesses $P_1,\ldots,P_{i-1}$ are already constructed. Then, the $i$th iteration is as follows:
\begin{itemize}
	\item First, we call $\candidateInd{W,r,k}$ with $W=P_1\cup P_2\cup \ldots \cup P_{i-1}$ consisting of the union of all the witnesses found so far. If this call returns $\bot$, then we can safely report that there is no distance-$r$ independent set of size $k$ in $G$. Otherwise, the output of the call becomes the next candidate $\tup a_i$.
	\item Second, we call $\witnessInd{\cal L,r,p}$, where $\cal L\coloneqq \{\tup a_1,\ldots,\tup a_i\}$ is the set of all the candidates found so far. If this call returns $\bot$, we report that there exists a distance-$r$ independent set of size $k$ in $G$. Otherwise, the witness output by the call becomes the next witness $P_i$, with which we can proceed to the next iteration.
\end{itemize}
Note that the correctness of the algorithm is not obvious: When the algorithm reports in the second point that there exists a distance-$r$ independent set of size $k$, then a priori there is no reason to assume that this is correct. In particular, the algorithm only provides this conclusion without providing any actual independent set. However, Fabia\'nski et al.~\cite{FabianskiPST19,Fabianski2018arxiv} proved that in classes of bounded expansion, the algorithm always finishes within a bounded number of rounds and provides a correct answer, provided the parameter $p$ is set large enough.

\begin{theorem}[consequence of {\cite[Theorems~5, Theorem~12, and Corollary~16]{Fabianski2018arxiv}}]\label{thm:ladder-bound}
	Let $\Cc$ be a graph class of bounded expansion and $r,k\in \N$. Then, there exist constants $p,\ell\in \Oh_{\Cc,r,k}(1)$ such that the ladder algorithm (\cref{alg:2}) run on any graph $G\in \Cc$ for parameters $r,k,p$, terminates within at most $\ell$ rounds and always outputs the correct answer. 
\end{theorem}

Thus, within a single run of the ladder algorithm as above, every $\witnessInd{\cal L,r,p}$ query involves a family $\cal L$ consisting of at most $\ell$ candidates, each being a $k$-tuple of vertices, and every $\candidateInd{W,r,k}$ query involves a vertex set $W$ with $|W|\leq p\ell$. We remark that similarly to \cref{thm:termination}, \cref{thm:ladder-bound} also holds in the larger generality where $\Cc$ is only assumed to be nowhere dense.

We now show how to implement the $\witnessInd{\cal L,r,p}$ and $\candidateInd{W,r,k}$ queries assuming access to the following generalizations of $\far{S,r}$ and $\near{S,r}$:
\begin{itemize}
	\item $\far{S,\rho}$: Given a set $S$ and a mapping $\rho\colon S\to \N$, return a vertex $v$ of $G$ such that $\dist_G(v,s)\geq\rho(s)$ for all $s\in S$; or $\bot$ if no such vertex exists.
	\item $\near{S,\rho}$: Given a set $S$ and a mapping $\rho\colon S\to \N$, return a vertex $v$ of $G$ such that $\dist_G(v,s)\leq \rho(s)$ for all $s\in S$; or $\bot$ if no such vertex exists.
\end{itemize}
In the following, for a function $\rho\colon S\to \N$ we denote $\max \rho\coloneqq \max_{s\in S} \rho(s)$. The reader should think that in the queries above we always have $\max \rho \leq r+1$ for the distance parameter $r$ we are working with. The generalization allows us to assign a different relevant distance $\rho(s)\leq r+1$ to every vertex $s$ of $S$.

\begin{lemma}\label{lem:candidateInd-far}
	Let $G$ be a graph, $W$ be a subset of vertices of $G$, and $k,r\in \N$ be parameters. Then the query $\candidateInd{W,r,k}$ can be answered in time $(r+2)^{\Oh(k|W|)}$ by performing at most $(r+2)^{k|W|}$ queries of the form $\far{S,\rho}$, where $S$ is a subset of $W$ and $\max \rho\leq r+1$.
\end{lemma}
\begin{proof}
	For a vertex $u$ of $G$, we define the \emph{profile} of $u$ as the function $\mathsf{prof}[u]\colon W\to \{0,1,\ldots,r,r+1\}$ such that for every $s\in W$,
	\[\mathsf{prof}[u](s)\coloneqq \begin{cases} \dist_G(u,s) &\textrm{if }\dist_G(u,s)\leq r,\\
		r+1 &\textrm{otherwise.}\end{cases}\]
	Let $\cal P\coloneqq \{0,1,\ldots,r,r+1\}^W$ be the set of all possible profiles. Note that $|\cal P|=(r+2)^{|W|}$. 
	
	Observe that for a $k$-tuple $\tup a$ of vertices, the $k$-tuple of profiles $(\mathsf{prof}[a_i]\colon i\in [k])$ determines whether $W$ is a distance-$r$ dependence witness for $\tup a$. Therefore, there is a set $\cal I\subseteq (\cal P)^k$ of $k$-tuples of profiles such that $W$ is a distance-$r$ dependence witness for $\tup a$ if and only if $(\mathsf{prof}[a_i]\colon i\in [k])\in \cal I$. Note that $\cal I$ can be computed in time $(r+2)^{\Oh(k|W|)}$ by investigating every $k$-tuple of profiles in $(\cal P)^k$ and deciding in time $(k|W|)^{\Oh(1)}$ whether it should be included in $\cal I$.
	
	Now, for each $\tup \pi=(\pi_1,\ldots,\pi_k)\in \cal I$ and each $i\in [k]$, we call $\far{S,\pi_i}$ to find, if existent, any vertex $a_i$ whose profile $\mathsf{prof}[a_i]$ is coordinate-wise not smaller than $\pi_i$. If for any $\tup \pi\in \cal I$ we manage to find all such vertices $a_i$, then we have found a tuple $\tup a=(a_1,\ldots,a_k)$ for which $W$ is not a distance-$r$ dependence witness; so $\tup a$ can be reported. And if this check fails for every $\tup \pi\in \cal I$, then we may safely conclude that $W$ is a distance-$r$ dependence witness for every $k$-tuple of vertices of $G$.
\end{proof}


\begin{lemma}\label{lem:witnessInd-near}
	Let $G$ be a graph, $p,k,r\in \N$, and $\cal L$ be a set of $k$-tuples of vertices of $G$. Then the query $\witnessInd{W,r,p}$ can be answered in time $(r+2)^{\Oh(pk|\cal L|)}$ by performing at most $(r+2)^{pk|\cal L|}$ queries of the form $\near{S,\rho}$, where $|S|\leq k|\cal L|$ and $\max \rho\leq r$.
\end{lemma}
\begin{proof}
	Let $L$ be the set of all the vertices featured in the tuples of $\cal L$. Note that $|L|\leq k|\cal L|$. We will use a slightly different notion of a profile than in the proof of \cref{lem:candidateInd-far}. For a vertex $u$ of $G$, its \emph{profile} is the function $\mathsf{prof}[u]\colon L\to \{0,1,\ldots,r,+\infty\}$ defined as follows: for $s\in L$,
	\[\mathsf{prof}[u](s)\coloneqq \begin{cases} \dist_G(u,s) &\textrm{if }\dist_G(u,s)\leq r,\\
		+\infty &\textrm{otherwise.}\end{cases}\]
	Let $\cal Q\coloneqq \{0,1,\ldots,r,+\infty\}^L$ be the set of all possible profiles. Note that $|\cal Q|=(r+2)^{|L|}\leq (r+2)^{k|\cal L|}$. 
	
	Observe that for a $p$-tuple of vertices $\tup b=(b_1,\ldots,b_p)$, the $p$-tuple of profiles $(\mathsf{prof}[b_i]\colon i\in [p])$ determines whether $\{b_1,\ldots,b_p\}$ is a distance-$r$ dependence witness for each of the $k$-tuples in $\cal L$. Therefore, there is a set $\cal J\subseteq (\cal Q)^p$ of $p$-tuples of profiles such that $\{b_1,\ldots,b_p\}$ is a distance-$r$ dependence witness for every $k$-tuple in $\cal L$ if and only if $(\mathsf{prof}[b_i]\colon i\in [p])\in \cal J$. Note that $\cal J$ can be computed in time $(r+2)^{\Oh(pk|\cal L|)}$ by investigating every $p$-tuple of profiles in $(\cal Q)^p$ and deciding in time $(pk|\cal L|)^{\Oh(1)}$ whether it should be included in $\cal J$.
	
	Now, for each $\tup \pi=(\pi_1,\ldots,\pi_p)\in \cal J$ and each $i\in [p]$, we let $L'\coloneqq \{s\in L~\mid~\pi_i(s)\textrm{ is finite}\}$ and call $\near{L',\pi_i|_{L'}}$ to find, if existent, any vertex $b_i$ whose profile $\mathsf{prof}[b_i]$ is coordinate-wise not larger than $\pi_i$. If for any $\tup \pi\in \cal J$ we manage to find all such vertices $b_i$, then $\{b_1,\ldots,b_p\}$ is a distance-$r$ dependence witness for all the tuples in $\cal L$; so $\{b_1,\ldots,b_p\}$ can be reported. And if this check fails for every $\tup \pi\in \cal J$, then we may safely conclude that there is no set of size at most $p$ that is a distance-$r$ dependence witness for all the tuples in $\cal L$.
\end{proof}

We conclude this section with a statement analogous to \cref{lem:summary-dom} that summarizes the ingredients needed to design a data structure for the \dris{r} problem.


\begin{lemma}\label{lem:summary-ind}
	Let $r,k\in \N$, $\Cc$ be a graph class of bounded expansion, and $p,\ell\in \N$ be the constants provided by \cref{thm:ladder-bound} for $\Cc,r,k$. Let $G$ be a dynamic graph that belongs to $\Cc$ at all times. Suppose that there are data structures $\nearVertexDS{G}{r,k\ell}$, supporting queries $\near{S,\rho}$ in $G$ with $|S|\leq k\ell$ and $\max \rho\leq r$, and $\farVertexDS{G}{r+1,p\ell}$, supporting queries $\far{S,\rho}$ in $G$ with $|S|\leq p\ell$ and $\max \rho\leq r+1$, so that $\nearVertexDS{G}{r,k\ell}$ and $\farVertexDS{G}{r+1,p\ell}$ have amortized update/query time $T$, initialization time $I$, and space complexity $M$. Then there is a data structure $\independentSetDS{G}{r,k}$ for $G$ that supports the~query
	\begin{itemize}
		\item $\is{}$: Decide whether in $G$ there exists a distance-$r$ independent set of size $k$. 
	\end{itemize}
	The amortized update time of $\independentSetDS{G}{r,k}$ is $(r+2)^{\Oh(pk\ell)}\cdot T$, the query time is $\Oh(1)$, the initialization time is $\Oh(I)$, and the space complexity is $\Oh(M)$.
	
	Moreover, we may allow the data structures $\nearVertexDS{G}{r,k\ell}$ and $\farVertexDS{G}{r+1,p\ell}$ to be randomized with error probability at most $\eps$, for any fixed parameter $\eps>0$ and against an oblivious adversary, with amortized update/query time, initialization time, and space complexity becoming $T\cdot \log \tfrac{1}{\eps}$, $I\cdot \log \tfrac{1}{\eps}$, and $M\cdot \log \tfrac{1}{\eps}$, respectively. In this case,  $\independentSetDS{G}{r,k}$ is also randomized with error probability at most $\eps$ against an oblivious adversary. The amortized update time becomes $(r+2)^{\Oh(pk\ell)}\cdot T\cdot \log\tfrac{1}{\eps}$, the initialization time becomes $(r+2)^{\Oh(pk\ell)}\cdot I\cdot \log \tfrac{1}{\eps}$, and the space complexity becomes  $(r+2)^{\Oh(pk\ell)}\cdot M\cdot \log \tfrac{1}{\eps}$.
\end{lemma}
\begin{proof}
	The proof is completely analogous to that of \cref{lem:summary-dom}, with \cref{thm:ladder-bound} used to argue that the ladder algorithm employed with parameter $p$ is correct and performs at most $\ell$ iterations, and \cref{lem:candidateInd-far,lem:witnessInd-near} used to reduce the at most $2\ell$ queries of the form $\witnessInd{\cal L,r,p}$ and $\candidateInd{W,r,k}$ to $(r+2)^{\Oh(pk\ell)}$ queries of the form $\near{S,\rho}$ and $\far{S,\rho}$ that are relayed to the data structures $\nearVertexDS{G}{r,k\ell}$ and $\farVertexDS{G}{r+1,p\ell}$. %We leave the details to the reader.
\end{proof}

\begin{comment}
\subsection{Old text}
In this work, we use the above framework to implement efficient
dynamic algorithms for the $r$-dominating set problem.
We explore the fact that we have the bound on the number of rounds
of Algorithm~\ref{alg:1} at hand. What remains to show is how
to efficiently answer $\candidate{W,r,k}$ and $\notDominated{S}$
queries in a dynamically changing graph. For that, we introduce
two dynamic data structures that are the building blocks for the above.

\begin{definition}
The $\nearVertexDS{G}{k,d}$ data structure maintains a dynamic graph $G$ by allowing the following operations:
\begin{itemize}
 \item $\init{G}$ - initializes with a graph $G$ on $n$ vertices
 \item $\add{uv}$ - inserts an edge $uv$ to $G$
 \item $\remove{uv}$ - removes an edge $uv$ from $G$
 \item $\near{S}$ - returns a vertex at distance at most $r$ from all vertices in $S$ if such vertex exists, otherwise it returns $\bot$.
\end{itemize}
\end{definition}

\begin{definition}
The $\farVertexDS{G}{k,d}$ data structure maintains a dynamic graph $G$ by allowing the following operations:
\begin{itemize}
 \item $\init{G}$ - initializes with a graph $G$ on $n$ vertices
 \item $\add{uv}$ - inserts an edge $uv$ to $G$
 \item $\remove{uv}$ - removes an edge $uv$ from $G$
 \item $\far{S}$ - returns a vertex at distance at least $r+1$ from
  $S$ if such vertex exists, otherwise it returns $\bot$.
\end{itemize}
\end{definition}


\begin{theorem}
Using $\nearVertexDS{G}{k,d}$ and $\farVertexDS{G}{k,d}$ data
structures,
one can implement Algorithm~\ref{alg:1} in time
$\sli(B(\delta^k_r,G)) \cdot ( k^{\sli(B(\delta^k_r,G))+1}
\mathsf{Q_{nv}} +
2k \cdot \mathsf{U_{fv}} + \mathsf{Q_{fv}})$, where
$\mathsf{Q_{nv}}$ is the query time
of $\nearVertexDS{G}{k,d}$ and $\mathsf{U_{fv}}$ and
$\mathsf{Q_{fv}}$ is the update
and the query time of
$\farVertexDS{G}{k,d}$ respectively.
\end{theorem}
\begin{proof}We show how to implement procedures $\candidate{W,r,k}$ and $\notDominated{S}$.
\paragraph*{Candidate.}We set up and maintain a
$\nearVertexDS{G}{k,d}$ data structure on the
input dynamic graph $G$. We implement query $\candidate{W,r,k}$
as follows.
We enumerate all
the partitions of vertices in $W=W_1 \cup \ldots \cup W_k$
into $k$ subsets. There is at most $k^{|W|}$ such partitions.
For each
such partition and each $i \in [k]$ we issue $v_i=\near{W_i}$
query, that returns a vertex at distance at most $r$ from
all vertices in $W_i$. If for some partition we can find near
vertices for all $i \in [k]$, then the set
$S=\{v_1,\ldots,v_k \}$ is a feasible candidate and we return
it. Otherwise, we return $\bot$. Thus, the total
number of queries issued to $\nearVertexDS{G}{k,d}$ is
$k^{|W|+1}$.

\paragraph*{Witness.} We set up and maintain a
$\farVertexDS{G}{k,d}$ data
structure on the input dynamic graph $G$ and some auxiliary
vertex $q \notin V(G)$. Each time a query
$\notDominated{S}$ is used by Algorithm~\ref{alg:1},
we issue $\far{S}$ query, which returns a vertex $v$ at
distance at least $r+1$ from $S$. Thus, $v$ is
a valid witness that $S$ is not a feasible solution, or, in
other words $\delta^k_r(S,v)$ does not hold.
We return a vertex $v$ if it exists,
or $\bot$ otherwise. This takes time proportional
to a query operation on $\farVertexDS{G}{k,d}$ data structure.

\end{proof}

The first step to use the above theorem is to bound the ladder index
$\sli(B(\delta_r^k,G))$ for specific classes of graphs.
After~\cite{FabianskiPST19} we get the following result.

\begin{theorem}[Theorem 11 and Lemma 5 in ~\cite{FabianskiPST19}]
For any $r,k \in \N$ and a nowhere dense class $\mathcal{C}$ of graphs
it holds that
$\sli(B(\delta_r^k,G)) \in k^{\Oh_{\mathcal{C},r}(1)}$ for any
$G \in \mathcal{C}$.
\end{theorem}

In Section~\ref{sec:distone} we develop the
following data structures for degenerate graphs.

\begin{theorem}\label{thm:distone-near}
Assume $n\ge 2$ and $d\ge 1$.
There exists
$\nearVertexDS{G}{k,d}$ data structure which
on a dynamic $d$-degenerate graph $G$ achieves the
following running times:
\begin{itemize}
\item $\init{G}$:  $\Oh(nd\cdot 2^{4d}(d^3+\log^3 n))$ time;
\item $\add{uv}$ and $\remove{uv}$:
$\Oh(2^{4d}(d^3+\log^3 n))$ amortized time;
\item $\near{S}$: $\Oh(|S| + d|S|^2\log n)$ time.
\end{itemize}
The structure uses $\Oh(nd\cdot 2^{4d})$ words of space.
\end{theorem}

\begin{theorem}\label{thm:distone-far}
Assume $n\ge 2$, $d\ge 1$, and let $\eps \in (0,1/2)$.
There exists a randomized data structure $\farVertexDS{G}{k,d}$ that,
against an oblivious adversary, maintains a dynamic $d$-degenerate
graph $G$ achieving the following running times:
\begin{itemize}
\item $\init{G}$:
$\Oh\!\left(nd\cdot 2^{4d}(d^3+\log^3 n)\log n \log \frac{1}{\eps}\right)$;
\item $\add{uv}$ and $\remove{uv}$: amortized
$\Oh\!\left(2^{4d}(d^3+\log^3 n)\log n \log \frac{1}{\eps}\right)$ time;
\item $\far{S}$: admits error probability bounded by $\epsilon$, takes time
\[
\Oh\!\left((|S|+2^{|S|} d {|S|}^2\log n)\log n \log \frac{1}{\eps}\right).
\]
\end{itemize}
The structure uses $\Oh\!\left(nd\cdot 2^{4d}\log n \log \frac{1}{\eps}\right)$ words of space.
\end{theorem}

By \cite{grohe_et_al:LIPIcs.FSTTCS.2013.21} bounded expansion graphs
are both nowhere dense and have bounded degeneracy. As a corollary, we
obtain the following result.


\begin{corollary}
Let $\mathcal{C}$ be a bounded expansion class of graphs.
There exists a randomized fully dynamic $\dominatingSetDS{n}$ data
structure against oblivious adversary,
that maintains an $n$-vertex graph $G \in \mathcal{C}$ by supporting
the following
operations:
\begin{itemize}
\item $\init{G}$: initializes the data structure in time
\item $\add{uv}$ and $\remove{uv}$:
\item $\ds{k}$:
\end{itemize}

TODO: details.
\end{corollary}

\azinline{i need section 5 to write the conclusions}

\begin{theorem}\label{thm:NearVertex}
There is a $\nearVertexDS{G}{k,d}$ data structure that maintains a dynamic graph $G$ of bounded
expansion, providing the following operations:
\begin{itemize}
 \item $\init{G}$ - initializes with a graph $G$ on $n$ vertices
 \item $\add{uv}$ - inserts an edge $uv$ to $G$
 \item $\remove{uv}$ - removes an edge $uv$ from $G$
 \item $\near{S}$ - returns a vertex at distance at most $r$ from all vertices in $S$ if such vertex exists, otherwise it returns $\bot$.
\end{itemize}
\end{theorem}
\begin{proof}Postponed to Section 5. \end{proof}

\begin{theorem}\label{thm:FarVertex}
There is a $\farVertexDS{G}{k,d}$ data structure that maintains a dynamic graph $G$ of bounded
expansion, providing the following operations:
\begin{itemize}
 \item $\init{G}$ - initializes with a graph $G$ on $n$ vertices
 \item $\add{uv}$ - inserts an edge $uv$ to $G$
 \item $\remove{uv}$- removes an edge $uv$ from $G$
 \item $\far{v}$ - returns a vertex at distance at least $r$ from $v$ if it exists,
 otherwise it returns $\bot$
\end{itemize}
\end{theorem}
\begin{proof}Postponed to Section 5. \end{proof}


\end{comment}
