\section{Preliminaries}\label{sec:prelims}

We write $\N$ for the set of nonnegative integers.
For a positive integer $k$, we denote $[k]\coloneqq \{1,\ldots,k\}$.
We follow the convention that if $\tup x$ is a tuple of objects, then the $i$th element of $\tup x$ is denoted by $x_i$.

\subsection{Graphs}

\paragraph{Basics.} We use standard graph terminology and notation. All graphs considered in this paper are finite, undirected, and simple (without loops or parallel edges), unless explicitly stated.

For a graph $G$, by $V(G)$ and $E(G)$ we denote the vertex set and the edge set of $G$, respectively. We also write $|G| \coloneqq |V(G)|$ and $\|G\| \coloneqq |E(G)|$. For vertices $u,v$ of $G$, by $\dist_G(u,v)$ we denote the distance between $u$ and $v$, defined as the smallest possible length of a path in $G$ with endpoints $u$ and $v$. This notation is extended to subsets naturally, e.g. $\dist_G(v,S)=\min_{s\in S}\dist_G(v,s)$. The \emph{closed neighborhood} of a vertex $u$ in $G$ is the set $N_G[u]$ that consists of $u$ and all the neighbors of $u$. By $\avgdeg(G)$ we denote the average degree in $G$; note that it is equal to $\frac{2\|G\|}{|G|}$.

A \emph{rooted graph} is a graph $G$ together with a tuple $\tup u$ of vertices of $G$ (not necessarily distinct), called the \emph{roots}. We denote it by $\rt{G}{\tup u}$. 

%Similarly, we may speak about rooted relational structures. The notions of homomorphisms, subgraphs, and induced subgraphs are naturally extended to rooted graphs or rooted induced subgraphs, by requiring that the source graph and the target graph have the same number of roots and that the $i$th root of the source graph is mapped to the $i$th root of the target graph, for every relevant $i$. In particular, if $\rt{H}{\tup x}$ and $\rt{G}{\tup u}$ are rooted graphs with $|\tup x|=|\tup u|$, then by $\Hom(\rt{H}{\tup x},\rt{G}{\tup u})$ we denote the set of all homomorphisms from $\rt{H}{\tup x}$ to $\rt{G}{\tup u}$; and similarly for subgraphs and induced subgraphs.

%We sometimes consider graphs with colored edges. In this case, $G$ is also equipped with a function $\mathrm{col} : V(G) \to C$, where $C$ is a set of admissible colors.\michal{Move to the place where we use it.}

We use the standard notion of directed graphs, again disallowing loops and parallel edges (edges with same head and tail). An \emph{oriented graph} is a directed graph $D$ where any pair of vertices can be the endpoints of at most one edge; that is, for distinct vertices $u,v$ we disallow that the edges $(u,v)$ and $(v,u)$ are simultaneously present. For an undirected graph $G$, an \emph{orientation} of $G$ is an oriented graph obtained from $G$ by choosing an orientation of every edge of $G$.

A directed graph is \textit{connected} if its underlying undirected graph is connected.

A graph $G$ is \emph{$d$-degenerate} if every subgraph of $G$ contains a vertex of degree at most $d$. The \emph{degeneracy} of $G$ is the smallest $d$ for which $G$ is $d$-degenerate. Note that if $G$ is $d$-degenerate, then it has an orientation with maximum outdegree at most $d$. Indeed, such an orientation can be obtained by iteratively removing from $G$ a vertex of the smallest degree (which is always at most~$d$) and orienting the edges incident to it away from it.

The \emph{arboricity} of a graph $G$ is the smallest $\alpha\in \N$ such that the edge of $G$ can be partitioned into $\alpha$ sets each of which forms a forest. It is well-known that if $d$ and $\alpha$ are the degeneracy and the arboricity of $G$, then we have
\[\alpha\leq d\leq 2\alpha.\]
In other words, degeneracy and arboricity of a graph are within a multiplicative factor of $2$, so they are essentially equivalent for our purposes. In this paper we choose to use degeneracy as a base parameter to state our results.

%\wninline{Będę chciał mieć definicję co znaczy ``data structure $S(G)$ for a dynamic graph $G$'', aby nie pisać tego 20 razy. Będę też chciał mieć warunek, że zawsze inicjalizujemy pustym grafem, aby się nie bawić w czasy inicjalizacji w każdym statemencie}

%\wninline{Definicja bnd expansion i nowhere-dense też się przyda. Chciałbym aby statementy o nowhere-dense używały $\delta$, a nie $\eps$, aby $\eps$ zostawić na prawdopodobieństwo błędu w randomizacji}

%\begin{definition}
%	Let $G$ be a graph and $v \in V(G), v' \not\in V(G)$. The operation of \textit{adding a pendant} $vv'$ produces a graph $G'$ such that $V(G') = V(G) \cup \{v'\}$ and $E(G') = E(G) \cup \{vv'\}$.
%\end{definition}


\paragraph*{Mappings between graphs.} For graphs $H,G$, a \emph{homomorphism} from $H$ to $G$ is a mapping $\varphi\colon V(H)\to V(G)$ such that for every pair of distinct vertices $u,v\in V(H)$, if $uv$ is an edge in $H$ then $\varphi(u)\varphi(v)$ is an edge in $G$. A \emph{subgraph isomorphism} is a homomorphism that is injective (i.e. $\varphi(u)\neq \varphi(v)$ for all distinct $u,v\in V(H)$), and an \emph{induced subgraph isomorphism} is a subgraph isomorphism where the implication stated above is in fact an equivalence: $uv$ is an edge in $H$ if and only if $\varphi(u)\varphi(v)$ is an edge in $G$. By $\Hom(H,G)$, $\Sub(H,G)$, and $\ISub(H,G)$ we denote the sets of all homomorphisms, subgraph isomorphisms, and induced subgraph isomorphisms from $H$ to $G$, respectively.

The definitions of homomorphisms and (induced) subgraphs isomorphisms can be naturally extended to rooted graphs, by requiring that the source graph $\rt{H}{\tup u}$ and the target graph $\rt{G}{\tup v}$ have the same number of roots ($|\tup u|=|\tup v|$) and that the $i$th root of the source graph is mapped to the $i$th root of the target graph ($\varphi(u_i)=v_i$, for all $i\in [|\tup u|]$). In particular, if $\rt{H}{\tup u}$ and $\rt{G}{\tup v}$ are rooted graphs with $|\tup u|=|\tup v|$, then by $\Hom(\rt{H}{\tup u},\rt{G}{\tup v})$ we denote the set of all homomorphisms from $\rt{H}{\tup u}$ to $\rt{G}{\tup v}$; and similarly for (induced) subgraph isomorphisms.


%\begin{definition}
%	We say that $\F$ is a \textit{type of mappings} if it is a function such that for undirected graphs $H, G$ with colored edges $\F(H, G)$ returns a subset of all mappings from $V(H)$ to $V(G)$.
%\end{definition}

%For example, homomorphisms, subgraph isomorphisms and induced subgraph isomorphisms are all types of mappings. We shall denote them by $\Hom$, $\Sub$ and $\ISub$, i.e. $\Hom(H, G)$, $\Sub(H, G)$ and $\ISub(H, G)$ will denote the set of all homomorphisms, subgraph isomorphisms and induced subgraph isomorphisms from $H$ to $G$, respectively.

\paragraph*{Relational structures.} We use the standard notion of relational structures considered in finite model theory. A \emph{signature} is a finite set $\Sigma$ of relation names, with each relation name $R\in \Sigma$ having a prescribed \emph{arity} $\ar(R)\in \N$. A \emph{$\Sigma$-structure} $\Af$ consists of a finite universe $U=U(\Af)$ and, for every relation name $R\in \Sigma$, its \emph{interpretation} $R^\Af\subseteq U^{\ar(R)}$. 

Here are some examples of modelling (variations of) graphs as relational structures, which will be used in this paper.
\begin{itemize}
	\item An undirected graph $G$ is modelled as a $\Sigma$-structure whose universe is the vertex set, and $\Sigma$ consists of a single binary (arity-$2$) relation that is always irreflexive and symmetric. To model directed graphs, we drop the symmetricity requirement.
	\item For a finite set of colors $C$, a \emph{$C$-colored graph} is an undirected graph $G$ together with a function $\mathrm{col}\colon E(G)\to C$ assigning every edge its color. We model $C$-colored graphs by having $|C|$ binary relations in the signature, where each relation selects edges of a different color. We may similarly define and model $C$-colored directed graphs.
\end{itemize}

Homomorphisms of relational structures are defined naturally: For two $\Sigma$-structures $\Af$ and $\Bf$, a homomorphism from $\Af$ to $\Bf$ is a mapping $\varphi\colon U(\Af)\to U(\Bf)$ such that for every $R\in \Sigma$ and $\tup u\in U(\Af)^{\ar(R)}$, that $\tup u\in R^{\Af}$ implies $\varphi(\tup u)\in R^{\Bf}$ (where $\varphi(\tup u)$ denotes coordinate-wise application). The notions of (induced) substructure isomorphisms then follow in the same way as for graphs, and so we may extend the $\Hom(\cdot,\cdot),\Sub(\cdot,\cdot),\ISub(\cdot,\cdot)$ notation to relational structures as well. 

Similarly to graphs, we may consider \emph{rooted relational structures} by equipping a structure $\Af$ with a tuple $\tup u$ of roots, which are (not necessarily distinct) elements of the universe; we denote it by $\rt{\Af}{\tup u}$. Homomorphisms and (induced) substructure isomorphisms can be again extended to rooted structures in the expected way. Thus we may speak about homomorphisms of colored directed graphs etc.

\paragraph*{Sparsity.} For $r\in \N$, we say that a \emph{depth-$r$ model} of a graph $H$ in a graph $G$ is a mapping $\eta$ from the vertices of $H$ to subsets of the vertices of $G$ satisfying the following two conditions:
\begin{itemize}
	\item The sets $\{\eta(u)\colon V(H)\}$ are pairwise disjoint and each of them induces a connected subgraph of $G$ of radius at most $r$.
	\item Whenever $uv$ is an edge in $H$, in $G$ there is an edge with one endpoint in $\eta(u)$ and the other in $\eta(v)$.
\end{itemize}
The sets $\{\eta(u)\colon u\in V(H)\}$ are called \emph{branch sets} of the model. We say that $G$ contains $H$ as a \emph{depth-$r$ minor} if $G$ contains a depth-$r$ model of $H$.

For a graph $G$ we define
\[\nabla_r(G)\coloneqq \max\{\,\avgdeg(H)~\colon~H\textrm{ is a depth-}r\textrm{ minor of }G\,\}.\]
For a graph class $\Cc$, we define
\[\nabla_r(\Cc)\coloneqq \sup_{G\in \Cc} \nabla_r(G).\]
Note that $\nabla_r(\Cc)$ may be equal to $+\infty$; this happens if graphs $G\in \Cc$ attain arbitrarily large values of $\nabla_r(G)$. This finite/infinite distinction underlies the central definition considered in this work.

\begin{definition}
	A class of graphs $\Cc$ has \emph{bounded expansion} if $\nabla_r(\Cc)$ is finite for every $r\in \N$. 
\end{definition}

Unpacking the definitions, this means that there exists a function $f\colon \N\to \N$ such that for every $G\in \Cc$ and a depth-$r$ minor $H$ of $G$, we have $\avgdeg(H)\leq f(r)$. Note that depth-$0$ minors are just subgraphs, hence every graph $G$ belonging to a class of bounded expansion $\Cc$ is $\nabla_0(\Cc)$-degenerate; this is a constant depending only on $\Cc$.

In our proofs, we will use stability of the notion of bounded expansion under two basic operations. The first is adding pendants.
For a graph $G$ and a vertex $v$ of $G$, the operation of \emph{adding a pendant} to $v$ produces a new graph $G'$ obtained from $G$ by adding a fresh vertex $v'$ and an edge $vv'$.
	
\begin{lemma} \label{obs:pendants-bnd-exp}
	Let $\Cc$ be a class of graphs of bounded expansion and let $\Cc$ be the class of all the graphs that can be obtained from a graph from $\Cc$ by repeatedly adding pendants (an arbitrary number times). Then $\Cc'$ is also of bounded expansion.
\end{lemma}
\begin{proof}
	Fix $r\in \N$.
	We prove that if $G'$ is obtained from $G$ by adding a pendant $v'$ to $v\in V(G)$, then
	\begin{equation}\label{eq:zubr}
		\nabla_r(G')\leq \max\left(1,\nabla_r(G)\right).
	\end{equation}
	By applying \eqref{eq:zubr} repeatedly to every graph in $G'\in \Cc'$, we conclude that
	\[\nabla_r(\Cc')\leq \max\left(1,\nabla_r(\Cc)\right),\]
	so $\Cc$ having bounded expansion implies that $\Cc'$ has bounded expansion as well.
	
	Towards \eqref{eq:zubr}, let $H'$ be a depth-$r$ minor of some $G'$, and let $\eta'$ be the witnessing model. We consider two cases. First, if $\eta'$ contains no branch set equal to $\{v'\}$, then it is easy to see that $\eta$ obtained by removing $v'$ from the branch set it is contained in (if any) is a depth-$r$ model of $H'$ in $G$. Hence $\avgdeg(H')\leq \nabla_r(\Cc)$. Second, if there exists $x\in V(H')$ such that $\eta'(x)=\{v'\}$, then $x$ must have degree at most $1$ in $H'$ and $\eta$ obtained from $\eta'$ by removing the branch set $\{v'\}$ is a depth-$r$ model of $H\coloneqq H'-x$ in $G$. It is easy to see that adding a pendant to a graph does not increase its average degree unless it stays below $1$, hence $\avgdeg(H')\leq \max(1,\avgdeg(H))\leq \max(1,\nabla_r(\Cc))$. This proves \eqref{eq:zubr} 
\end{proof}

The next operation is taking congested shallow minors. For $c,d\in \N$, a \emph{congestion-$c$ depth-$d$ model} $\eta$ of a graph $H$ in a graph $G$ is defined similarly to a (standard) depth-$d$ model, except for the following amendments:
\begin{itemize}
	\item In the first condition, we allow the branch sets $\{\eta(u)\colon u\in V(H)\}$ to intersect, but every vertex $v$ of $G$ may belong to at most $c$ of them.
	\item In the second condition, we demand that whenever $uv$ is an edge of $H$, the branch sets $\eta(u)$ and $\eta(v)$ intersect or there is an edge connecting them in $G$.
\end{itemize}
Note that for $c=1$, these are just standard depth-$d$ models. 

For a class of graphs $\Cc$, by $\Minors^{c,d}(\Cc)$ we denote the class of all congestion-$c$ depth-$d$ minors of graphs from $\Cc$. We will use the following statement.

\begin{theorem}[TODO]\label{thm:cong-minors}
	For every graph class $\Cc$ of bounded expansion and $c,d\in \N$, the class $\Minors^{c,d}(\Cc)$ also has bounded expansion.
\end{theorem}

We use \cref{thm:cong-minors} to argue a classic result from Sparsity about classic of bounded expansion admitting \emph{fraternal augmentations} of bounded maximum outdegree. Let $G$ be a graph and $\vec G$ be an orientation of $G$. The \emph{fraternal augmentation} of $\vec G$ is the undirected graph $G'$ obtained from $\vec G$ as follows:
\begin{itemize}
	\item for every pair of edges $(u,w),(v,w)\in E(\vec G)$ with a common tail $w$, add an undirected edge $uv$; and
	\item forget the orientation of all the edges of $\vec G$.
\end{itemize}
Note that thus, $G'$ is a supergraph of $G$, obtained by adding all the \emph{fraternal} edges $uv$ as above. The following lemma is the key observation about the behavior of fraternal augmentations on classes of bounded expansion.

\begin{lemma}\label{lem:fraternal}
	Let $\Cc$ be a graph class of bounded expansion and $d\in \N$. Define $\Fra(\Cc,d)$ to be the graph class constructed as follows: for every $G\in \Cc$ and orientation $\vec G$ of $G$ with maximum outdegree at most $d$, include in $\Fra(\Cc,d)$ the fraternal augmentation of $\vec G$. Then $\Fra(\Cc,d)$ also has bounded expansion.
\end{lemma}
\begin{proof}
	We observe that if $G'$ is the fraternal augmentation of an orientation $\vec G$ of $G$ of maximum outdegree~$d$, then $G'$ is a congestion-$(d+1)$ depth-$1$ minor of $G$. Indeed, if for $u\in V(G)$ we define
	\[\eta(u)\coloneqq \{u\}\cup \{v~\mid~(v,u)\in E(\vec G)\},\]
	then $\eta$ is a congestion-$(d+1)$ depth-$1$ model of $G'$ in $G$. It follows that $\Fra(\Cc,d)\subseteq \Minors^{d+1,1}(\Cc)$, so $\Fra(\Cc,d)$ has bounded expansion by \cref{thm:cong-minors}.
\end{proof}


%	 Let $G\in \Cc$ be such that $G'$ is obtained from $G$ by repeatedly adding pendants. 

%Let $H'$ be an $r$-shallow minor of some $G' \in \G'$. We have that $H'$ can be obtained from some $r$-shallow minor $H$ of some $G \in \G'$ by repeatedly adding pendants. As adding a pendant increases the number of vertices and edges by one, there exists $c \in \N$ such that $||H'|| - ||H|| = c = |H'| - |H|$, so $\frac{||H'||}{|H'|} = \frac{||H|| + c}{|H| + c} \le \max(1, \frac{||H||}{|H|}) \le \max(1, f_{\G}(r))$, hence we can set $f_{\G'}(r) \coloneqq \max(1, f_{\G}(r))$ as a function bounding densities of $r$-shallow minors of $\G'$, proving that $\G'$ is of bounded expansion.


%\begin{observation} \label{obs:pendants-nowhere-dense}
%	Let $\G$ be a nowhere-dense class of graphs and let $\G'$ be a class of graphs that can be obtained from $\G$ by adding pendants arbitrarily many times. Then $\G'$ is also nowhere-dense.
%\end{observation}

%\begin{proof}
%	Let $f_{\G} : \N \to \N$ be a function such that no $r$-shallow minor $H$ of any $G \in \G$ has a clique bigger than $f_{\G}(r)$.
	
%	Similarly as in \cref{obs:pendants-bnd-exp}, every $r$-shallow minor of graphs from $\G'$ can be obtained by repeatedly adding pendants to some $r$-shallow minor of a graph from $\G$. As no pendant can be a part of a clique with more than two vertices, we have that we can set $f_{\G'}(r) \coloneqq \max(2, f_{\G}(r))$ as a function bounding sizes of cliques in $r$-shallow minors of $\G'$, proving that $\G'$ is also nowhere-dense.
%\end{proof}


\subsection{Computation}

\paragraph*{Computation model.} When working with a graph on $n$ vertices in the algorithmic context, we use the standard word RAM model with words of length $\Oh(\log n)$. In particular, space complexity of data structures is measured in the number of words occupied. We remark that all the numbers appearing in the computation, e.g. vertex weights or homomorphism counts, will be always bounded polynomially in $n$, and hence they fit within a constant number of words and arithmetic operations on them can be executed in constant time.

\paragraph*{Dynamic graphs and data structures.} Let us introduce some terminology to facilitate speaking about fully dynamic data structures for graphs that are updated over time.

By a \emph{dynamic graph} we mean a graph $G$ whose vertex set stays fixed, but whose edge set is modified over time by updates of the following types:
\begin{itemize}
	\item \emph{insert} an edge $uv$, if not already present; and
	\item \emph{remove} the edge $uv$, if present.
\end{itemize}
By a \emph{data structure} for a dynamic graph $G$ we mean a data structure that maintains $G$ and supports a superset of the following basic methods:
\begin{itemize}
	\item $\init{n}$: Initialize the data structure on an edgeless graph $G$ on $n$ vertices.
	\item $\add{uv}$: Insert the edge $uv$.
	\item $\remove{uv}$: Remove the edge $uv$.
\end{itemize}
These methods will be typically augmented by some further queries, specified in the description of every data structure. By the \emph{update time} of a data structure we mean the time complexity of methods $\add{uv}$ and $\remove{uv}$, and by the \emph{initialization time} we mean the time complexity of method $\init{n}$.

In most of this work we will silently assume that every data structure for a dynamic graph $G$ maintains a dictionary of the edges of $G$, and hence we may efficiently (in logarithmic time) query whether two vertices of $G$ are adjacent, and get the corresponding edge. This is immaterial when the claimed complexity bounds are polylogarithmic with an unspecified degree of the polylogarithm. When we care about specific degrees of the polylogarithms, we explicitly describe the method of handling the edge set.

Some of our data structures are randomized \emph{against an oblivious adversary}. By this we mean that the data structure is run on some stream of updates and queries and the answer to every query is correct with probability lower bounded by a specified value (typically, $1-\tfrac{1}{n}$). In particular, the stream is fixed upfront and the consecutive updates/queries \emph{cannot} be altered depending on the data structure's answers. We remark that while the answer to every query is correct with probability at least, say, $1-\tfrac{1}{n}$, the answers to the queries are \emph{not} independent random variables. In fact, in all the randomized data structures proposed in this work, the answers to the queries depend on some common random bits fixed upon the initialization of the data structure.


\begin{theorem} \label{thm:maintaining-augmentation}
	Placeholder na stwierdzenie, że da się dynamicznie utrzymywać augmentacje. Zapewne chcemy zacytować Theorem 4 z \cite{DvorakTumaArxiv}.
\end{theorem}